### 2018-09-18 paper/Paper Title.Rmd
created a rough outline for the paper

---
title: "Paper Title (Something about Applying LIME to Bullet Data)"
author: "Katherine Goode and Heike Hoffman"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background/Literature Review

# Problem Statement

# Overview of LIME

# Explain the Data (TS 224)

# Results (using LIME on a TS)