### 2019-01-30 presentations/jsm/jsm abstract.Rmd
added rfscore datasets and drafted an abstract for JSM

---
title: "JSM Abstract"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
---

### To submit a session abstract, you will need the following:

#### Session Type (Contributed, Topic-Contributed*, Roundtable)

Contributed

#### Session Subtype (Paper, Speed, Poster, A.M. or P.M. Roundtable)

Paper, speed, or poster (*Which should I list as the perferred session?*)

#### Primary Sponsor (ASA Committee, ASA Section, JSM Sponsoring Society). A pull-down screen with a pre-approved list is provided.

(*I am not sure what to list for this?*)

#### Abstract Title

An Assessment of LIME Input Options when Interpreting Random Forest Predictions for Bullet Matching Data
(*This seems too long and wordy...*)

#### Abstract Text (1200 character limit)

Random forests are known for their accurate predictive abilities, but they are a part of the family of machine learning models that lack interpretability. Recently, a technique called LIME was developed to provide local interpretations for predictive models. The technique fits a linear regression model to perturbations created from the observed data in a localized region around a prediction of interest and the predicted values associated with the perturbations computed using the complex model. The coefficents from the linear model are used to determine the influential variables. We apply LIME to a random forest fit to a bullet matching dataset using the lime package currently available in R. The LIME algorithm requires several input options that affect the way that the perturbations are created and how the model is interpreted. We assess the current options avaiable in the package and propose some new methods.

(*What are your thoughts aobut this? Is the focus of this abstract apporpriate for the talk?*)

#### Abstract Keywords

Key Words: random forest, LIME, interpretability

(*Are these okay?*)

#### Presenting author and co-author contact information. Please note, there can only be one presenting author. Accurate and unique email addresses are required for all authors.

Presenting author: Katherine Goode <kgoode@iastate.edu>

Co-author: Heike Hoffman <hofmann@iastate.edu>

#### Note:

- If the abstract can be considered for alternate session subtypes. The program committee will not have space for all paper abstracts received. Acceptance is more likely if the presentation can be made in different session formats.
- Information on abstract submission from: http://ww2.amstat.org/meetings/jsm/2019/submissions.cfm