### 2019-02-09 papers/An Assessment of LIME Explanations from a Random Forest Model.Rmd
worked more on journals...

---
title: "An Assessment of LIME Explanations from a Random Forest Models"
author: "Katherine Goode and Heike Hofmann"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
# Set knitr options
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE, 
                      fig.pos = "center")
```