### 2019-02-12 journals/00-ideas.Rmd
added an ideas journal (00)

---
title: "00-ideas"
author: "Katherine Goode"
date: "2/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Concerns with LIME

- We think that the model explainer needs to be close enough to the complex model that it is trying to explain in order to do a good job of providing explanations. For example, the binned regression works okay with neural networks, but it does not work well with a random forest. Maybe a tree explainer or a logistic regression explainer would work better with a random forest.

# Ideas to Try

- Could try fitting a regression with interactions and see if LIME does a good job of explaining the model.

# General Thoughts about LIME

- LIME is kind of like a jacknife technique

# Current Plan

- Write a paper that shows that LIME does not work well with random forests.