### 2019-02-26 presentations/graphics group/applying lime to a random forest and the problems that ensue.Rmd
worked on pres

---
title: "Applying LIME to a Random Forest Model and the Problems that Ensue"
author: "Katherine Goode"
date: "March 1, 2019"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

# Get Started