### 2019-02-26 presentations/graphics group/applying lime to a random forest and the problems that ensue.Rmd
updates before meeting

---
title: "My Story of Applying LIME to a Random Forest Model and the Problems that Ensued"
author: "Katherine Goode"
date: "March 1, 2019"
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

## Overview

1. Overview of LIME
2. Hamby bullet matching data
3. Applying LIME to the bullet data
4. Issues with LIME
5. Attempts at a Solution
6. Future Work