### 2019-02-28 presentations/graphics group/An Application of LIME to a Random Forest.Rmd
worked more on presentation

---
title: "[![](./images/lindens.jpg)](https://en.wikipedia.org/wiki/Tilia) An Application of LIME to a Random Forest"
author: <font size="5"> Katherine Goode </font>
date: <font size="5"> ISU Graphics Group, March 1, 2019 </font>
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, eval = FALSE)
```

# Overview

.pull-left[
### Goals for the talk
- Explain the LIME algorithm
- Present an application of LIME
- Highlight problems with this application

### The plan

1. Explanation of LIME 
2. Hamby bullet data
3. Applying LIME to the bullet data
4. Issues with LIME
5. Attempts at a solution
6. Future work
]

.pull-right[
<br>
<br>
<img src="./images/lime_drawing.png" width = 400>
]

---

class: inverse, center, middle

# What is <span style="color:lime">LIME</span>?

---

# Motivation for LIME

<br>

[![](./images/blackbox.png)](https://en.wikipedia.org/wiki/Black_box)

### Black Box Prediction Models

- Offer great predictive ability 
- Loss of interpretability
- Difficult to assess trustworthiness

### Enter LIME

- **L**ocal **I**nterpretable **M**odel-Agnostic **E**xplanations
- Developed by computer scientists ([Ribeiro, Singh, and Guestrin](https://arxiv.org/pdf/1602.04938.pdf))
- Designed to assess if a black box predictive model is trustworthy
- Produces "explanations" for individual predictions

---

# Meaning of LIME

.pull-left[
### <span style="color:lime">L</span>ocal

- Focuses on behavior of a complex model at a local level

### <span style="color:lime">I</span>nterpretable

- Produces easily interpretable "explanations"

### <span style="color:lime">M</span>odel-Agnostic

- Works with any black-box predictor

### <span style="color:lime">E</span>xplanations

- Provides insight into individual predictions
]

.pull-right[
<br>
<br>
.center[<img src="./images/local.png" width=300>]
.center[<font size="4">Figure 3 in Ribeiro, Singh, and Guestrin (2016)</font>]
<br>
<br>
.center[<img src="./images/image_explanation.png" width=500>]
.center[<font size="4">Figure 4 in Ribeiro, Singh, and Guestrin (2016)</font>]
]

---

# An Example <font size="3">(from Ribeiro, Singh, and Guestrin (2016))</font>

.pull-left[
### 1. Black Box Model 
- Model predicts whether a patient has the flu
- Apply the model to a new patient
- Predicts that the patient has the flu
- Can this prediction be trusted?
]

.pull-right[
### 2. LIME
- Apply LIME to *this* case
- LIME returns the most important variables in *this* prediction
- Colors indicate
    + <span style="color:green">green</span>: evidence supporting the flu
    + <span style="color:red">red</span>: evidence against the flu
- Can this prediction be trusted?
]

.center[<img src="./images/example.png" width=700>]
.center[<font size="4">Figure 1 in Ribeiro, Singh, and Guestrin (2016)</font>]

---

# Idea of the LIME Procedure

Start with training data, complex model, and one prediction from the test data
1. Use the features in the training data to create a new dataset of many perturbations.
2. Obtain a prediction for each perturbation using the complex model.
3. Compute a similarity score between the test observation and each perturbation. 
4. Perform feature selection to choose $k$.
5. Fit a linear regression (weighted by the similarity scores) to the perturbations with the $k$ chosen features (standardized).
  $$\mbox{prediction} \sim \mbox{chosen feature 1} + \cdots + \mbox{chosen feature k}$$
6. Use the coefficients from the regression to "explain" the predictions.

| | Feature 1 | <span style="color:blue">Feature 2</span> | <span style="color:blue">Feature 3</span> | Feature 4 | Similarity Score | Prediction |
| --: | :--: | :--: | :--: | :--: | :--: | :--: |
| test observation | 2 | 4 | 2 | 10 | exact | $\hat{y}_{\mbox{obs}}$ |
| perturbation 1 | 2 | 3 | 2 | 12 | very close | $\hat{y}_1$ |
| $\vdots$ | $\vdots$ | ... | ... | ... | ... | 
| perturbation 5000  | 10 | 0 | 13 | 22 | not close | $\hat{y}_{5000}$ |

---

# Using LIME

### Implementations of LIME

- Developers created a Python package called [lime](https://github.com/marcotcr/lime)
- Thomas Pedersen created an R package also called [lime](https://github.com/thomasp85/lime)

.pull-left[
```{r eval = FALSE}
#install.packages("lime")
library(lime)
?lime
?explain
```
]

.pull-right[![](./images/lime.png)]
---

class: inverse, center, middle

# Hamby Bullet Matching Data

---
class: inverse, center, middle

# Applying LIME to the bullet data

---

# <span style="color:lime">LIME</span> Procedure with Iris Data (in R)

## Step 0: Split iris data into training and testing datasets

```{r, echo = TRUE}
# Iris dataset
iris[1:3, ]

# Split up the data set into training and testing datasets
iris_test <- iris[1:5, 1:4]
iris_train <- iris[-(1:5), 1:4]

# Create a vector with the responses for the training dataset
iris_lab <- iris[[5]][-(1:5)]
```

---

## Step 1: Fit a complex model to the training data

```{r, echo = TRUE}
# Create random forest model on iris data
library(caret)
rf_model <- train(iris_train, iris_lab, method = 'rf')

# Can use the complex model to make predictions
Pred <- predict(rf_model, iris_test)
Actual <- iris[1:5, 5]
data.frame(iris_test, Pred, Actual)
```

---

## Step 2: Obtain distributions of the variables from the training data

```{r, echo = TRUE}
# Create an explainer object
library(lime); explainer <- lime(iris_train, rf_model)

# Sepal length quantiles obtained from training data
explainer$bin_cuts$Sepal.Length

# Probability distribution for sepal length
explainer$feature_distribution$Sepal.Length
```

---

## | Histogram of Sepal Length from Training Data

```{r, fig.width = 8, fig.height = 5, fig.align = "center"}

# Histogram of sepal length from training data
ggplot(iris_train, aes(x = Sepal.Length)) +
  geom_histogram(bins = 10, color = "grey") +
  geom_vline(xintercept = explainer$bin_cuts[[1]][[1]], color = "green", size = 1) +
  geom_vline(xintercept = explainer$bin_cuts[[1]][[2]], color = "green", size = 1) +
  geom_vline(xintercept = explainer$bin_cuts[[1]][[3]], color = "green", size = 1) +
  geom_vline(xintercept = explainer$bin_cuts[[1]][[4]], color = "green", size = 1) +
  geom_vline(xintercept = explainer$bin_cuts[[1]][[5]], color = "green", size = 1) +
  labs(x = "Sepal Length", y = "Count")
```

---

## Step 3: Sample from each of the variable distributions $n$ times

Histograms of predictor variables from training data

```{r, out.width = "500px", fig.align = "center"}
knitr::include_graphics("./Graphics/training_distributions.png")
```

---

## | Histograms of $n=5000$ samples from distributions of training variables for each of the five testing cases

```{r, out.width = "550px", fig.align = "center"}
knitr::include_graphics("./Graphics/samples.png")
```

---

## Step 4: Obtain predictions for sampled values using the complex model

* For each testing case use the random forest model to make a prediction for each of the $n=5000$ samples
* In the iris data, the predictions are represented by the probability that a flower is a particular species

```{r}
perturb <- read.csv("./Data/perturb_predictions.csv")
perturb
```

---

## Step 5: Obtain similarity score between data observation and sampled values

We need to determine how similar a sampled case is to the observed case in the testing data

Case 1 from testing data:
```{r}
iris_test[1, ]
```

First sample from training data variable distributions associated with case 1 of testing data:
```{r}
perturb[1, 1:4]
```

----

LIME uses exponential kernel function
  $$\pi_{x_{obs}}(x_{sampled}) = exp\left\{\frac{−D(x_{obs}, \ x_{sampled})^2}{σ^2}\right\}$$
where

$x_{obs}$: observed data vector to predict

$x_{sampled}$: sampled data vector from distribution of training variables

$D(\cdot \ , \ \cdot)$: distance function such as euclidean distance, cosine distance, etc.  
$\sigma$: width (default set to 0.75 in `lime`)

---

## Step 6: Perform feature selection by fiting a model to the sampled data and associated predictions (weighted by the similarity scores)

* The user can specify the number of variables (features) they would like to select: $m$
* With the iris data, the following three models will be fit to perform variable selection to select $m=2$ features:

$$\mbox{P(setosa)} \sim \mbox{Sepal.Length} + \mbox{Sepal.Width} + \mbox{Petal.Length} + \mbox{Petal.Width}$$
$$\mbox{P(versicolor)} \sim \mbox{Sepal.Length} + \mbox{Sepal.Width} + \mbox{Petal.Length} + \mbox{Petal.Width}$$
$$\mbox{P(virginica)} \sim \mbox{Sepal.Length} + \mbox{Sepal.Width} + \mbox{Petal.Length} + \mbox{Petal.Width}$$

----

* To perform variable selection `lime` supports: 
    - forward selection with ridge regression
    - highest weight with ridge regression
    - LASSO
    - tree models
    - auto: forward selection if $m\le6$, highest weight otherwise

---

## Step 7: Fit a simple model to regress the predictions on the $m$ selected predictor variables (weighted by the similarity scores)

* Currently, `lime` is programmed to use ridge regression as the "simple" model
* If the response is categorical, the user can select how many categories they want to explain
* In this example, only setosa will be explained
* If petal length and sepal length were selected as the most important features for the first case in the testing data, then the simple model is

  $$\mbox{P(Setosa)} \sim \mbox{Petal.Length} + \mbox{Sepal.Length} $$

---

## Step 8: Extract the feature weights and use them as the explanations

* Steps 3 through 8 are all performed using the `explain` function in `lime`
* Steps 3 through 7 all happen behind the scenes
* The "feature weights" that are extracted are the coefficients for the predictor variables from the ridge regression

```{r, echo = TRUE}
# Explain new observation
explanation <- explain(iris_test, explainer, n_labels = 1, 
                       n_features = 2, n_permutations = 5000,
                       feature_select = 'auto')
```

----

```{r, echo = TRUE}
explanation[1:2, 1:6]
explanation[1:2, 7:10]
explanation[1:2, 11:13]
```

---

## | Plot of explanations for predictions from random forest model trained on iris data

```{r, echo = TRUE}
plot_features(explanation)
```

---
class: inverse, center, middle

# Issues with LIME

---
class: inverse, center, middle

# Attempts at a Solution

---
class: inverse, center, middle

# Future Work