### 2019-03-01 presentations/graphics group/An Application of LIME to a Random Forest.Rmd

---
title: "[![](./images/lindens.jpg)](https://en.wikipedia.org/wiki/Tilia) An Application of LIME to a Random Forest"
author: <font size="5"> Katherine Goode </font>
date: <font size="5"> ISU Graphics Group, March 1, 2019 </font>
output:
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include = FALSE}
options(htmltools.dir.version = FALSE, eval = FALSE)

# Load libraries
library(gt)
library(randomForest)
library(tidyverse)
```

# Overview

.pull-left[
### The plan

1. Explanation of LIME 
2. Hamby bullet data
3. Applying LIME to the random forest
4. Issues with LIME and attempts at a solution
6. Future work
]

.pull-right[
<br>
<br>
<img src="./images/lime_drawing.png" width = 400>
]

---

class: inverse, center, middle

# What is <span style="color:lime">LIME</span>?

---

# Motivation for LIME

<br>

[![](./images/blackbox.png)](https://en.wikipedia.org/wiki/Black_box)

### Black Box Prediction Models

- Offer great predictive ability 
- Loss of interpretability
- Difficult to assess trustworthiness

### Enter LIME

- **L**ocal **I**nterpretable **M**odel-Agnostic **E**xplanations
- Developed by computer scientists ([Ribeiro, Singh, and Guestrin](https://arxiv.org/pdf/1602.04938.pdf))
- Designed to assess if a black box predictive model is trustworthy
- Produces "explanations" for individual predictions

---

# Meaning of LIME

.pull-left[
### <span style="color:lime">L</span>ocal

- Focuses on behavior of a complex model at a local level

### <span style="color:lime">I</span>nterpretable

- Produces easily interpretable "explanations"

### <span style="color:lime">M</span>odel-Agnostic

- Works with any black-box predictor

### <span style="color:lime">E</span>xplanations

- Provides insight into individual predictions
]

.pull-right[
<br>
<br>
.center[<img src="./images/local.png" width=300>]
.center[<font size="4">Figure 3 in Ribeiro, Singh, and Guestrin (2016)</font>]
<br>
<br>
.center[<img src="./images/image_explanation.png" width=500>]
.center[<font size="4">Figure 4 in Ribeiro, Singh, and Guestrin (2016)</font>]
]

---

# An Example <font size="3">(from Ribeiro, Singh, and Guestrin (2016))</font>

.pull-left[
### 1. Black Box Model 
- Model predicts whether a patient has the flu
- Apply the model to a new patient
- Predicts that the patient has the flu
- Can this prediction be trusted?
]

.pull-right[
### 2. LIME
- Apply LIME to *this* case
- LIME returns the most important variables in *this* prediction
- Colors indicate
    + <span style="color:green">green</span>: evidence supporting the flu
    + <span style="color:red">red</span>: evidence against the flu
- Can this prediction be trusted?
]

.center[<img src="./images/example.png" width=700>]
.center[<font size="4">Figure 1 in Ribeiro, Singh, and Guestrin (2016)</font>]

---

# General LIME Procedure

Start with training data, complex model, and one prediction from the test data
1. Create perturbations based on training data.
2. Obtain a prediction for each perturbation using the complex model.
3. Compute a similarity score between the test observation and each perturbation. 
4. Perform feature selection to choose $k$.
5. Fit a linear regression (weighted by the similarity scores) to the perturbations with the $k$ chosen features (standardized).
  $$\mbox{prediction} \sim \mbox{chosen feature 1} + \cdots + \mbox{chosen feature k}$$
6. Use the coefficients from the regression to "explain" the predictions.

| | Feature 1 | <span style="color:blue">Feature 2</span> | <span style="color:blue">Feature 3</span> | Feature 4 | Similarity Score | <span style="color:blue">Prediction</span>|
| --: | :--: | :--: | :--: | :--: | :--: | :--: |
| test observation | 2 | 4 | 2 | 10 | exact | $\hat{y}_{\mbox{obs}}$ |
| perturbation 1 | 2 | 3 | 2 | 12 | very close | $\hat{y}_1$ |
| $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | 
| perturbation 5000  | 10 | 0 | 13 | 22 | not close | $\hat{y}_{5000}$ |

---

# Using LIME

### Implementations of LIME

- Developers created a Python package called [lime](https://github.com/marcotcr/lime)
- Thomas Pedersen created an R package also called [lime](https://github.com/thomasp85/lime)

![](./images/lime.png)

### Key Functions in the lime R package

```{r eval = FALSE}
#install.packages("lime")
library(lime)
?lime
?explain
```

---

class: inverse, center, middle

# Hamby Bullet Matching Data

---

# Hamby Bullet Study [<font size="3">James E. Hamby et. al. (2009)</font>](https://cdn2.hubspot.net/hub/71705/file-15668427-pdf/docs/aftespringvol41no2pages99-110.pdf)

- Sets of bullets from both “known” and “unknown” gun barrels were sent to firearm examiners around the world
- 240 total sets created
    - 10 barrels in a set
    - 35 bullets in a set 
        - 20 known (2 from the same barrel)
        - 15 unknowns (at least 1 and no more than 3 from each barrel)
- Examiners asked to use the known bullets to identify which barrels the unknown bullets were fired from

.center[<img src="./images/hamby.png" width=350>]
.center[<font size="4">Results table from Hamby et. al. (2009)</font>]

---

# Automated Bullet Matching Algorithm

- CSAFE has access to some of the bullet sets
- [Hare, Hofmann, and Carriquiry ](https://projecteuclid.org/euclid.aoas/1514430288)(2017) develop an automated algorithm to determine whether two bullets are a match
- High definition scans of bullets used to obtain signatures associated with each land
- Developed variables that measure how similar two signatures are
- Fit a random forest model to predict whether two bullets were a match

.center[<img src="./images/signatures.png" width=575>]
.center[<font size="4">Figure from Hare, Hofmann, and Carriquiry (2017)</font>]

---

# Training Data

#### Sets 173 and 252

```{r}
hamby173and252_train <- read.csv("./data/hamby173and252_train.csv") # 83,028 rows
```

#### Variables Identifying the Bullets

```{r}
hamby173and252_train %>% select(1:8) %>% slice(1:5)
```

---

# Training Data

#### Signature Similarity Features

```{r}
hamby173and252_train %>% select(9:17) %>% slice(1:5) %>% round(2)
```

#### Response and RF Score

```{r}
hamby173and252_train %>% select(18:19) %>% slice(1:5)
```

---

# Random Forest Model

The random forest model fit to the training data in Hare, Hofmann, and Carriquiry can be found in the bulletr R package. It is called `rtrees`.

.pull-left[
```{r}
# Load the bulletr library
library(bulletr)
```
```{r}
# Importance values from 
# the random forest
rtrees$importance
```
]

.pull-right[
```{r}
# Number of trees
rtrees$ntree

# Number of variables randomly 
# sampled at each split
rtrees$mtry

# Confusion matrix
rtrees$confusion
```
]

---

# Testing Data

```{r}
hamby224_test <- read.csv("./data/hamby224_test.csv")
glimpse(hamby224_test)
```

---

# Testing Data

#### Results from Applying `rtrees` to Hamby 224 (Sets 1 and 11)

<br>

```{r echo = FALSE}
data.frame(Truth = hamby224_test$samesource %>% na.omit(),
           RF_Prediction = as.logical(predict(rtrees, hamby224_test %>%
                                       select(ccf:sum_peaks) %>% 
                                       na.omit()))) %>%
  count(Truth, RF_Prediction) %>%
  rename(Count = n) %>%
  gt::gt()
```

<br>

- Want to know which were the driving variables for individual predictions
- Especially great when the model is wrong
- Variable importance scores provide global interpretations
- We want local interpretations

---

class: inverse, center, middle

# Applying <span style="color:lime">LIME</span> to the Random Forest

---

# Step 0: Starting Objects

**Training Data**: `hamby173and252_train`

**Complex Model**: `rtrees`

**Case of Interest**: case 1 from `hamby224_test`

<br>

```{r echo = FALSE}
hamby224_test %>% filter(case == 1) %>% select(case:land2) %>% gt()
```

<br>

```{r echo = FALSE}
hamby224_test %>% filter(case == 1) %>% select(ccf:rfscore) %>% round(2) %>% gt()
```

---

# Step 1: Create Perturbations

#### Estimate the feature distributions

.pull-left[
- quantile bins (default with 4 bins)
- equally spaced bins
]
.pull-right[
- normal approximation 
- kernel density estimation
]
.center[
```{r echo = FALSE, fig.height = 4.5, fig.width = 9}
limeout <- lime::lime(x = hamby173and252_train %>% select(ccf:sum_peaks), model = rtrees)
ggplot(hamby173and252_train, aes(x = ccf)) + 
  geom_histogram(bins = 30) + 
  geom_vline(xintercept = limeout$bin_cuts$ccf[[1]], color = "green", size = 1) +
  geom_vline(xintercept = limeout$bin_cuts$ccf[[2]], color = "green", size = 1) +
  geom_vline(xintercept = limeout$bin_cuts$ccf[[3]], color = "green", size = 1) +
  geom_vline(xintercept = limeout$bin_cuts$ccf[[4]], color = "green", size = 1) +
  geom_vline(xintercept = limeout$bin_cuts$ccf[[5]], color = "green", size = 1) +
  geom_vline(xintercept = hamby224_test %>% 
               filter(case == 1) %>% 
               select(ccf) %>%
               pull(), color = "blue", size = 1) +
  labs(x = "CCF", y = "Frequency") +
  theme(text = element_text(size = 18))
```
]
.center[<font size="4">Histogram of CCF from Training Data with LIME Quantile Bins (blue line is ccf value of case 1)</font>]

---

# Step 1: Create Perturbations

#### Draw many samples from the estimated feature distributions to obtain the perturbations

- If the a binning estimation method is used, numeric features are converted to indicator variables.
    - 1 if in same bin as case 1
    - 0 o.w.

| | ccf same bin? | rough_cor same bin? | $\cdots$ | sum_peaks same bin?|
| --: | :--: | :--: | :--: | :--: | 
| case 1 | 1 | 1 | $\cdots$ | 1 |
| perturbation 1 | 0 | 1 | $\cdots$ | 1 |
| $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ |
| perturbation 5000  | 0 | 0 | $\cdots$ | 1|

- Otherwise, features are left as is.

---

# Step 2: Obtain Predictions

#### Use the complex model to obtain a prediction for each perturbation

- Use the `rtrees` model to do this

<br>

| | ccf same bin? | rough_cor same bin? | $\cdots$ | sum_peaks same bin?| prediction |
| --: | :--: | :--: | :--: | :--: | :--: |
| case 1 | 1 | 1 | $\cdots$ | 1 | not a match |
| perturbation 1 | 0 | 1 | $\cdots$ | 1 | match |
| $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ |
| perturbation 5000  | 0 | 0 | $\cdots$ | 1 | not a match |

---

# Step 3: Similarity Score

#### Compute a similiarity score between the test observation and each perturbation.

- Default method in R package is gower distance
    - between 0 and 1 
    - 0 indicates exactly the same
- Other distance metrics can be specified

| | ccf same bin? | rough_cor same bin? | $\cdots$ | sum_peaks same bin?| distance | prediction |
| --: | :--: | :--: | :--: | :--: | :--: |
| case 1 | 1 | 1 | $\cdots$ | 1 | 0 | not a match |
| perturbation 1 | 0 | 1 | $\cdots$ | 1 | 0.1 | match |
| $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ | $\vdots$ |
| perturbation 5000  | 0 | 0 | $\cdots$ | 1 | 0.8 | not a match |

---

# Step 4: Feature Selection

#### Perform feature selection to choose $k$ features.

- Can specify the number of features to select
    - We have been choosing 3 features
- Options in lime
     - auto: forward selection if $m\le6$, highest weight otherwise (default)
     - forward selection with ridge regression
     - highest weight with ridge regression
     - LASSO
     - tree models
- We have been using the default method

$$\mbox{rf score}=\beta_0 + \beta_1 I[\mbox{ccf in same bin}] + \beta_2 I[\mbox{rough_cor in same bin}] +\cdots + \beta_9 I[\mbox{sum_peaks in same bin}]$$

---

# Step 5: Simple Model

#### Fit a linear regression (weighted by the similarity scores) to the perturbations with the 3 chosen features.

- Features are standardized
- Currently, lime uses ridge regression as the "simple" model
- If the response is categorical, the user can select how many categories they want to explain
    - We have only been explaining "TRUE"
- Example model:

$$\mbox{rf score}=\beta_0 + \beta_1 I[\mbox{ccf in same bin}] + \beta_2 I[\mbox{cms in same bin}] + \beta_3 I[\mbox{matches in same bin}]$$

---

# Step 6: Interpret

#### Use the coefficients from the regression to "explain" the predictions.

- Since features are standardized, we can compare the coefficients
- Largest $\beta_i$ is conisdered the most important

---

# R Code

## lime Function

The `lime` function estimates the feature distributions (first part of step 1).

```{r}
set.seed(20190229)
out_lime <- lime::lime(x = hamby173and252_train %>% select(ccf:sum_peaks), 
                      model = lime::as_classifier(rtrees))
out_lime$bin_cuts$ccf
out_lime$feature_distribution$ccf
```

---

# R Code

## explain Function

The `explain` function performs the rest of the steps.

```{r}
out_explain <- lime::explain(x = hamby224_test %>% filter(case == 1) %>%
                               select(ccf:sum_peaks),
                             explainer = out_lime,
                             labels = TRUE,
                             n_features = 3)
out_explain[,1:6]
out_explain[,7:10]
```

---

# R Code

## R Package Plot of Explanations

The function `plot_features` let you visualize the explanations.

```{r fig.width = 10, fig.height = 5}
lime::plot_features(out_explain)
```

---

# Shiny App

<br>

.center[<img src="./images/app.png" width=1100>]

---
class: inverse, center, middle

# Issues with <span style="color:lime">LIME</span> and Attempts at a Solution

---

# Initial Concerns

We first applied lime using all default settings
- explanations did not make sense
- $R^2$ values were very low
- predictions from the simple model were poor

.center[<img src="./images/explain1.png" width=450>]
.center[<font size="4">Example where the random forest prediction is wrong. The explanations are from from 4 quantile bins.</font>]

---

# More Issues

Then we tried more of the feature distribution estimaton methods
- explanations still not great
- $R^2$ values still very low
- predictions from the simple model still poor
- explanations dependent on the input

.center[<img src="./images/explain2.png" width=450>]
.center[<font size="4">This is the same case with explanations are from from 3 equally spaced bins.</font>]

---

# New Binning Method

Next we created our own way to select bins using trees

```{r echo = FALSE, message = FALSE}
source("../../code/helper_functions.R")
```

.pull-left[
```{r}
# Example using treebink
treebink(
  y = hamby173and252_train$samesource, 
  x = hamby173and252_train$ccf, 
  k = 5
)
```
]
.pull-right[
.center[<img src="./images/tree.png" height = 500>]
]
---

# Comparing Methods

#### Feature distribution estimaton methods:

- 2 to 5 quntile bins
- 2 to 5 equally spaced bins
- 2 to 5 tree based bins with `samesoure` as the response
- 2 to 5 tree based bins with `rfscore` as the response
- normal approximation
- kernel density estimation

#### Comparison metrics:

- MSEs (rf prediction vs simple model predictions)
- $R^2$
- Consistency across number of bins
- Consistency across 10 reps

---

# Comparing MSEs

.center[<img src="./images/mse.png" width = 900>]

---

# Comparing $R^2$ Values

.center[<img src="./images/r2.png" width = 900>]

---

# Comparing Consistency Across Bins

.center[<img src="./images/firstfeature.png" width = 1100>]

---

# Comparing Consistency Across Bins

.center[<img src="./images/secondfeature.png" width = 1100>]

---

# Comparing Consistency Across Bins

.center[<img src="./images/thirdfeature.png" width = 1100>]

---

# Comparing Consistency Across Reps

.center[<img src="./images/reps1.png" width = 1100>]

---

# Comparing Consistency Across Reps

.center[<img src="./images/reps2.png" width = 1100>]

---

# Comparing Consistency Across Reps

.center[<img src="./images/reps3.png" width = 1100>]

---

# Conclusions

#### Key Observations

- equally spaced bins are dependent on the nubmer of bins
- tree based bins usually have the lowest MSEs
- rfscore tree based bins usually have the highest $R^2$ for the binning options
- all options results in low $R^2$ values
    - we think that the linear regression is not similar enough to a random forest to produce good explanations
- explanations are relatively consistent across reps for first and second features
- non-binning methods produce very different results

---
class: inverse, center, middle

# Future Work