### 2019-03-11 journals/08-iris_comparison/08-iris_comparison.Rmd
started working on iris comparison

---
title: "08-iris_comparison"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE)
```

I am interested in how lime performs on other random forest models. This journal applies all of the input lime options to a random forest model fit to the iris data.

```{r}
# Load libraries
library(caret)
library(furrr)
library(future)
library(gretchenalbrecht)

# Source functions
source("../../code/helper_functions.R")
```

# Training and Testing Data

```{r}
# Set a seed
set.seed(20190311)

# Randomly select two cases from within each of the three species of irises
selected <- c(sample(1:50, 4), sample(51:100, 4), sample(101:151, 4))

# Split up the features of the data into training and testing parts
iris_train <- iris[-selected, ]
iris_test <- iris[selected, ] %>% mutate(case = selected)
```

# Random Forest Model

```{r}
# Random forest model run on the iris data
iris_model <- train(iris_train %>% select(-Species), iris_train %>% pull(Species), method = 'rf')

# Predictions made using the random forest model on the testing data
iris_model_predict <- predict(iris_model, iris_test)

# Matrix of observed and predicted values
iris_test_obs_pred <- matrix(c(iris_test %>% pull(Species), iris_model_predict), ncol = 2, byrow = FALSE)

# Specify the column names
colnames(iris_test_obs_pred) <- c("Observed", "Predicted")

# Print the table
knitr::kable(iris_test_obs_pred, align = 'c')
```

# Apply LIME

```{r}
# Apply the run_lime function if the lime results file does not already exist
if(!file.exists("./data/iris_lime_inputs.rds")) {

  # Specify the input options to use with lime
  iris_lime_inputs <- list(bin_continuous = c(rep(TRUE, 15), 
                                                  rep(FALSE, 2)),
                               quantile_bins = c(rep(TRUE, 5), 
                                                 rep(FALSE, 5), 
                                                 rep(TRUE, 7)),
                               nbins = c(rep(2:6, 3), 
                                         rep(4, 2)),
                               use_density = c(rep(TRUE, 15), TRUE, FALSE),
                               bin_method = c(rep("quantile_bins", 5),
                                              rep("equally_spaced", 5),
                                              rep("tree", 5),
                                              rep("quantile_bins", 2)),
                               response = c(rep(NA, 10), 
                                            rep("Species", 5), 
                                            rep(NA, 2)))

  # Tell R to run the upcoming code in parallel
  plan(multiprocess)

  # Apply lime to the full training data with the specified input options
  iris_lime_explain <- future_pmap(.l = iris_lime_inputs,
             .f = run_lime, # run_lime is one of my helper functions
             features = names(iris %>% select(-Species)),
             train = iris_train,
             test = iris_test %>% select(-case, -Species),
             rfmodel = iris_model,
             label = "virginica",
             nfeatures = 3,
             seed = TRUE)

  # Separate the lime and explain function results from the full data
  iris_lime <- map(iris_lime_explain, function(list) list$lime)
  iris_explain <- map_df(iris_lime_explain, function(list) list$explain)

  # Name the items in the lime list
  names(iris_lime) <- map_chr(1:17, function(case) 
      sprintf("case: bin_continuous = %s, quantile_bins = %s, nbins = %0.f, use_density = %s, bin_method = %s, response = %s",
              iris_lime_inputs$bin_continuous[case],
              iris_lime_inputs$quantile_bins[case],
              iris_lime_inputs$nbins[case],
              iris_lime_inputs$use_density[case],
              iris_lime_inputs$bin_method[case],
              iris_lime_inputs$response[case]))

  # Turn the lime input options into a dataframe before saving it
  iris_lime_inputs <- iris_lime_inputs %>%
    unlist() %>%
    matrix(ncol = length(iris_lime_inputs), 
           dimnames = list(NULL, names(iris_lime_inputs))) %>%
    as.data.frame() %>%
    mutate(case = 1:length(iris_lime_inputs$quantile_bins)) %>%
    select(case, bin_continuous:response)

  # Save the lime objects
  saveRDS(iris_lime_inputs, "./data/iris_lime_inputs.rds")
  saveRDS(iris_lime, "./data/iris_lime.rds")
  saveRDS(iris_explain, "./data/iris_explain.rds")

} else {

  # Load in the lime objects
  iris_lime_inputs <- readRDS("./data/iris_lime_inputs.rds")
  iris_lime <- readRDS("./data/iris_lime.rds")
  iris_explain <- readRDS("./data/iris_explain.rds")

}
```

```{r}
# Create the test_explain combined data if the file does not already exist
if(!file.exists("./data/iris_test_explain.rds")) {

  # Join the data and the explanations and edit and add additional variables
  iris_test_explain <- iris_test %>%
    mutate(case = as.character(case),
           rfscore = predict(iris_model, iris_test, type = "prob") %>% pull(virginica)) %>%
    full_join(iris_explain, by = "case") %>%
    mutate(case = factor(case),
           feature_desc = factor(feature_desc),
           feature = factor(feature),
           nbins = factor(nbins)) %>%
    arrange(nbins) %>%
    mutate(nbins = as.numeric(as.character(nbins)),
           situation = ifelse(bin_continuous == TRUE & bin_method == "quantile_bins", 
                              sprintf("%.0f quantile", nbins),
                              ifelse(bin_continuous == TRUE & bin_method == "equally_spaced",
                                     sprintf("%.0f equally spaced", nbins),
                                     ifelse(bin_continuous == TRUE & bin_method == "tree" &
                                              response == "Species",
                                            sprintf("%.0f tree", nbins),
                                            ifelse(bin_continuous == FALSE & 
                                                   use_density == TRUE, 
                                                   "kernel density", 
                                                   "normal approximation")))) %>%
             fct_relevel("2 quantile", "3 quantile", "4 quantile",
                         "5 quantile", "6 quantile", "2 equally spaced",
                         "3 equally spaced", "4 equally spaced",
                         "5 equally spaced", "6 equally spaced",
                         "2 Species tree", "3 Species tree",
                         "4 Species tree", "5 Species tree",
                         "6 Species tree")) %>%
    mutate(bin_situation = ifelse(bin_method == "quantile_bins" & 
                                  bin_continuous == TRUE,
                                  "quantile",
                                  ifelse(bin_method == "equally_spaced" & 
                                         bin_continuous == TRUE,
                                         "equally spaced", 
                                         ifelse(bin_method == "tree" & 
                                                bin_continuous == TRUE & 
                                                response == "Species",
                                                "Species tree", 
                                                ifelse(bin_continuous == FALSE & 
                                                       use_density == TRUE, 
                                                       "kernel density", 
                                                       "normal approximation"))))) %>%
    mutate(bin_situation = factor(bin_situation))

  # Save the combined test and explain data
  saveRDS(iris_test_explain, "./data/iris_test_explain.rds")

} else {

  # Load in the data
  iris_test_explain <- readRDS("./data/iris_test_explain.rds")

}
```

This code creates and saves a dataset with summaries of the explanations.

```{r}
# Create the lime comparison data if the file does not already exist
if(!file.exists("./data/iris_lime_comparisons.rds")) {

  # Create a data frame with the interesting information relating to the different
  # evaluations of lime and compute the difference and mean between the rf and rr
  # model predictions
  iris_lime_comparisons <- iris_test_explain %>%
    select(-data, -prediction) %>%
    group_by(case, bin_continuous, quantile_bins, nbins, use_density, bin_method) %>%
    slice(1) %>%
    ungroup() %>%
    mutate(diff = rfscore - model_prediction,
           mean = (rfscore + model_prediction) / 2)

  # Save the lime comparison data frame
  saveRDS(iris_lime_comparisons, "./data/iris_lime_comparisons.rds")

} else {

  # Load in the lime comparison data frame
  iris_lime_comparisons <- readRDS("./data/iris_lime_comparisons.rds")

}
```

# Visualizing LIME Results

```{r}
# Summarizing lime results
iris_lime_results <- iris_lime_comparisons %>%
  group_by(situation, bin_situation, bin_continuous, quantile_bins, nbins, use_density, 
           bin_method, response) %>%
  summarise(mse = (sum(diff^2)) / length(diff),
            ave_r2 = mean(model_r2)) %>%
  ungroup()
```

```{r}
# Plot of MSEs
iris_lime_results %>%
  mutate(bins = ifelse(nbins == 4 & bin_continuous == FALSE, "other", nbins)) %>%
  ggplot(aes(x = bin_situation, y = mse, color = bin_situation)) +
  geom_point() + 
  facet_grid(~ bins, scales = "free_x", space = "free_x") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), 
        legend.position = "none") + 
  labs(x = "Bin Type", y = "MSE", color = "") + 
  scale_color_gretchenalbrecht(palette = "winged_spill", discrete = TRUE)
```

```{r}
# Plot of R^2s
iris_lime_results %>%
  mutate(bins = ifelse(nbins == 4 & bin_continuous == FALSE, "other", nbins)) %>%
  ggplot(aes(x = bin_situation, y = ave_r2, color = bin_situation)) +
  geom_point() + 
  facet_grid( ~ bins, scales = "free_x", space = "free_x") + 
  theme_bw() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") + 
  labs(x = "Bin Type", y = "Average R2", color = "") + 
  scale_color_gretchenalbrecht(palette = "winged_spill", discrete = TRUE)
```

```{r}
# Creates a dataset that creates columns of the chosen features by lime from first to third
iris_chosen_features <- iris_test_explain %>%
  filter(!(bin_situation %in% c("kernel density", "normal approximation"))) %>%
  filter(!is.na(rfscore)) %>%
  select(situation, bin_situation, nbins, case, feature, feature_weight) %>%
  mutate(feature_weight_abs = abs(feature_weight)) %>%
  arrange(situation, case, desc(feature_weight_abs)) %>%
  mutate(explainer = rep(c("first", "second", "third"), length(situation) / 3)) %>%
  select(-feature_weight, -feature_weight_abs) %>%
  spread(explainer, feature)
```

```{r}
# Computes and plots the values of kappa
iris_chosen_features %>%
  ungroup() %>%
  select(-situation) %>%
  gather(key = chosen, value = feature, first:third) %>%
  spread(nbins, feature) %>%
  select(-case) %>%
  rename(bins2 = "2", bins3 = "3", bins4 = "4", bins5 = "5", bins6 = "6") %>%
  group_by(bin_situation, chosen) %>%
  summarise(kappa = 
              irr::kappam.fleiss(matrix(c(bins2, bins3, bins4, bins5, bins6), 
                                        ncol = 5))$value) %>%
  arrange(chosen, bin_situation) %>%
  ggplot(aes(x = bin_situation, y = kappa, color = chosen, group = chosen)) +
  geom_point() + 
  geom_line() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  scale_color_gretchenalbrecht(palette = "pink_cloud", discrete = TRUE) +
  labs(x = "Bin Type", y = "Kappa", color = "Feature")
```

```{r}
# Plots the top feature for each case for each bin situation
iris_chosen_features %>%
  group_by(bin_situation, case) %>%
  mutate(nlevels = length(levels(factor(first)))) %>%
  ungroup() %>%
  arrange(bin_situation, desc(nlevels), case) %>%
  mutate(order = rep(rep(1:length(unique(case)), each = 5, 3))) %>%
  ggplot(aes(x = nbins, y = order, fill = first)) + 
  geom_tile() + 
  facet_grid( ~ bin_situation, scales = "free", space = "free_y") + 
  theme_bw() + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) + 
  labs(x = "Number of Bins", y = "Test Case", fill = "Feature", title = "First Feature")

# Plots the second feature for each case for each bin situation
iris_chosen_features %>%
  group_by(bin_situation, case) %>%
  mutate(nlevels = length(levels(factor(second)))) %>%
  ungroup() %>%
  arrange(bin_situation, desc(nlevels), case) %>%
  mutate(order = rep(rep(1:length(unique(case)), each = 5, 3))) %>%
  ggplot(aes(x = nbins, y = order, fill = second)) + 
  geom_tile() + 
  facet_grid( ~ bin_situation, scales = "free", space = "free_y") + 
  theme_bw() + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) + 
  labs(x = "Number of Bins", y = "Test Case", fill = "Feature", title = "Second Feature")

# Plots the third feature for each case for each bin situation
iris_chosen_features %>%
  group_by(bin_situation, case) %>%
  mutate(nlevels = length(levels(factor(third)))) %>%
  ungroup() %>%
  arrange(bin_situation, desc(nlevels), case) %>%
  mutate(order = rep(rep(1:length(unique(case)), each = 5, 3))) %>%
  ggplot(aes(x = nbins, y = order, fill = third)) + 
  geom_tile() + 
  facet_grid( ~ bin_situation, scales = "free", space = "free_y") + 
  theme_bw() + 
  theme(axis.text.y = element_blank(),
        axis.ticks.y = element_blank()) + 
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) + 
  labs(x = "Number of Bins", y = "Test Case", fill = "Feature", title = "Third Feature")
```

# Session Info

```{r}
sessionInfo()
```