### 2019-06-12 journals/09-understanding_lime/09-understanding_lime.Rmd
adding journals back

---
title: "09-understanding_lime"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE, 
                      cache = TRUE)
```

# Overview

```{r}
library(furrr)
library(future)
library(lime)
library(randomForest)
library(tidyverse)

# Source functions
source("../../code/helper_functions.R")
```

# The Data

The cleaned training and testing datasets are loaded in below.

```{r}
# Load in the training data (Hamby Data 173 and 252)
hamby173and252_train <- read.csv("../../data/hamby173and252_train.csv")

# Load in the testing data (Hamby Data 224 Sets 1 and 11)
hamby224_test <- read.csv("../../data/hamby224_test.csv")
```

A vector containing the features used in the random forest `rtrees` is created below that will be used when creating some of the plots in this journal.

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features_ordered <- data.frame(feature = rownames(bulletr::rtrees$importance), 
                                  MeanDecreaseGini = bulletr::rtrees$importance) %>%
  arrange(desc(MeanDecreaseGini))

top2_features <- rf_features_ordered$feature[1:2]
```

# Random Forest Model

```{r}
# Fit a random forest model with the top two most important features in the model
if(!file.exists("./data/rfmodel.rds")) {

  # Fit the random forest  
  rfmodel <- randomForest(factor(samesource) ~ rough_cor + ccf, data = hamby173and252_train)

  # Save the random forest
  saveRDS(rfmodel, "./data/rfmodel.rds")

} else {

  # Load the rf model
  rfmodel <- readRDS("./data/rfmodel.rds")
}
```

```{r}
# Create a dataset with the test data input to the random forest model 
# and the predictions from the random forest model
hamby224_test_subset <- hamby224_test %>%
  na.omit() %>%
  select(case:land2, rough_cor, ccf, samesource) %>%
  mutate(rfpred = predict(rfmodel, 
                           newdata = hamby224_test %>% select(rough_cor, ccf) %>% na.omit()),
         rfscore = predict(rfmodel, 
                           newdata = hamby224_test %>% select(rough_cor, ccf) %>% na.omit(),
                           type = "prob")[,2])
```

# Applying LIME

The functions `lime` and `explain` from the lime package are applied below to the random forest model.

```{r}
# Create or load the lime and explain objects for the main effects logisitic regression model
if(!file.exists("./data/lime_explain_sub.rds")) {

  # Set a seed
  set.seed(20190315)

  # Apply lime
  lime_sub <- lime(x = hamby173and252_train %>% select(rough_cor, ccf), 
                   model = as_classifier(rfmodel))

  # Apply explain
  explain_sub <- explain(x = hamby224_test %>%
                             select(rough_cor, ccf) %>% 
                             na.omit(), 
                           explainer = lime_sub, 
                           n_labels = 1,
                           n_features = 2)

  # Join the lime and explain objects in a list
  lime_explain_sub <- list(lime = lime_sub, explain = explain_sub)

  # Save the lime and explain objects
  saveRDS(lime_explain_sub, "./data/lime_explain_sub.rds")

} else {

  # Load the lime and explain objects
  lime_explain_sub <- readRDS("./data/lime_explain_sub.rds")

}
```

```{r}
create_bin_df <- function(bin_cut) {
 data.frame(quantile = c("0%", "25%", "50%", "75%", "100%"), 
            bin_cut) 
}

bin_cuts <- purrr::map_df(lime_explain_sub$lime$bin_cuts, 
                          create_bin_df,
                          .id = "feature") %>%
  spread(key = feature, value = bin_cut)

bin_cuts
```

# Visualizations of Data, Model, and LIME

```{r}
(case1 <- lime_explain_sub$explain %>% filter(case == 1))
```

```{r}
ggplot() +
  geom_point(data = hamby173and252_train %>% filter(samesource == FALSE),
             aes(x = rough_cor, y = ccf),
             color = "darkgrey",
             alpha = 0.2) +
  geom_point(data = hamby173and252_train %>% filter(samesource == TRUE),
             aes(x = rough_cor, y = ccf),
             color = "darkorange") + 
  geom_vline(xintercept = bin_cuts$rough_cor) + 
  geom_hline(yintercept = bin_cuts$ccf) +
  theme_minimal()
```

```{r}
ggplot() + 
  geom_point(data = hamby224_test_subset, 
             aes(x = rough_cor, y = ccf, color = rfscore),
             alpha = 0.5) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
  geom_point(data = lime_explain_sub$explain %>% 
               filter(case == 1) %>%
               select(feature, feature_value) %>%
               spread(key = feature, value = feature_value),
             aes(x = rough_cor, y = ccf)) +
  geom_vline(xintercept = bin_cuts$rough_cor) + 
  geom_hline(yintercept = bin_cuts$ccf) + 
  theme_minimal()
```

# Session Info

```{r}
sessionInfo()
```