### 2019-08-09 papers/lime_diagnostics/code/figure-creation.Rmd
added lime concept figures - worked on intro - started lime procedure

---
title: "Figure Creation"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

# Overview

This document contains the code for creating figures to be included in the LIME diagnostic paper.

```{r}
# Load packages
library(cowplot)
library(gower)
library(graphics)
library(tidyverse)
```

# Conceptual Depiction of LIME

### Good Fit

```{r}
# Generate data
set.seed(20190624)
lime_data <- data.frame(feature1 = sort(runif(250, 0, 1)),
                      feature2 = sample(x = 1:250, size = 250, replace = FALSE)) %>%
  mutate(prediction = if_else(feature1 >= 0 & feature1 < 0.1, (0.3 * feature1) + rnorm(n(), 0, 0.01), 
                      if_else(feature1 >= 0.1 & feature1 < 0.3, rbeta(n(), 1, 0.5), 
                      if_else(feature1 >= 0.3 & feature1 < 0.5, sin(pi* feature1) + rnorm(n(), 0, 0.5),
                      if_else(feature1 >= 0.5 & feature1 < 0.8, -(sin(pi* feature1) + rnorm(n(), 0, 0.1)) + 1,
                      if_else(feature1 >= 0.8 & feature1 < 0.9, 0.5 + runif(n(), -0.5, 0.5), 
                              0.5 + rnorm(n(), 0, 0.3)))))))

# Specify a prediction of interest
prediction_of_interest <- data.frame(feature1 = 0.07,
                                feature2 = 200,
                                prediction = 0.05, 
                                color = factor("Prediction of Interest"))

# Compute the distance between the prediction of interest 
lime_data$distance <- (1 - gower_dist(x = prediction_of_interest, y = lime_data))^20

# Fit the interpretable explainer model
lime_explainer <- lm(prediction ~ feature1 + feature2, data = lime_data, weights = distance)

# Predictions versus feature 1
plot_feature1 <- ggplot(lime_data, aes(x = feature1, y = prediction, size = distance)) + 
  geom_abline(intercept = coef(lime_explainer)[[1]] + coef(lime_explainer)[[3]]*prediction_of_interest$feature2, 
              slope = coef(lime_explainer)[[2]], 
              size = 1) +
  geom_point() + 
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature1, y = prediction),
             color = "#EFB084",
             size = 5) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 1", 
       y = "Black-Box Prediction") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")

# Predictions versus feature 2
plot_feature2 <- ggplot(lime_data, aes(x = feature2, y = prediction, size = distance)) +
  geom_abline(intercept = coef(lime_explainer)[[1]] + coef(lime_explainer)[[2]]*prediction_of_interest$feature1,
              slope = coef(lime_explainer)[[3]],
              size = 1) +
  geom_point() +
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature2, y = prediction),
             color = "#EFB084",
             size = 5) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 2",
       y = "Black-Box Prediction") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")

# Join the two feature plots
lime_plots <- plot_grid(plot_feature1, plot_feature2, ncol = 1)

# Create a plot to extract a legend for the prediction of interest
lime_legend_plot <- ggplot(prediction_of_interest, aes(x = feature1, y = prediction, color = color)) +
  geom_point(size = 5) +
  scale_color_manual(values = c("#EFB084")) +
  theme_classic(base_family = "Times") +
  labs(color = "") +
  theme(legend.position = "bottom")

# Extract the legend
lime_legend <- get_legend(lime_legend_plot)

# Join the title, plots, legend, and caption into one figure
lime_concept <- plot_grid(lime_plots, lime_legend,
                         ncol = 1,
                         rel_heights = c(0.9, 0.1))

# Preview the plot
lime_concept

# Save the plot
ggsave(filename = "../figures/lime_concept.png",
       plot = lime_concept,
       height = 6, width = 4)
```

### Poor Fit

```{r}
# Compute the distance between the prediction of interest 
lime_data$distance_bad <- (1 - gower_dist(x = prediction_of_interest, y = lime_data))^1

# Fit the interpretable explainer model
lime_explainer_bad <- lm(prediction ~ feature1 + feature2, data = lime_data, weights = distance_bad)

# Predictions versus feature 1
plot_feature1_bad <- ggplot(lime_data, aes(x = feature1, y = prediction, size = distance_bad)) + 
  geom_point() + 
  geom_abline(intercept = coef(lime_explainer_bad)[[1]] + coef(lime_explainer_bad)[[3]]*prediction_of_interest$feature2, 
              slope = coef(lime_explainer_bad)[[2]], 
              size = 1) +
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature1, y = prediction),
             color = "#EFB084",
             size = 6) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 1", 
       y = "Black-Box Prediction") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")

# Predictions versus feature 2
plot_feature2_bad <- ggplot(lime_data, aes(x = feature2, y = prediction, size = distance_bad)) + 
  geom_point() + 
  geom_abline(intercept = coef(lime_explainer_bad)[[1]] + coef(lime_explainer_bad)[[2]]*prediction_of_interest$feature1, 
              slope = coef(lime_explainer_bad)[[3]], 
              size = 1) +
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature2, y = prediction),
             color = "#EFB084",
             size = 6) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 2", 
       y = "Black-Box Prediction") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")

# Join the two bad feature plots
lime_plots_bad <- plot_grid(plot_feature1_bad, plot_feature2_bad, ncol = 1)

# Join the title, plots, legend, and caption into one figure
lime_figure_bad <- plot_grid(lime_plots_bad, lime_legend,
                         ncol = 1, 
                         rel_heights = c(0.9, 0.1))

# View the figure
lime_figure_bad

# Save the plot
ggsave(filename = "../figures/lime_figure_bad.png",
       plot = lime_figure_bad,
       height = 6, width = 4)
```