### 2020-09-17 README.Rmd
added files for comparing random forest models to rtrees

---
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LIME Research <img align="right" height="200" src="./figures/lime.png">

Repository for research projects on LIME

Link to csafe server: https://isu-csafe.stat.iastate.edu/rstudio/

## Ideas

**LIME diagnostics**

- turn repetitive code into functions
- try out different weights
- change distance matrix in seriation to be 5-# of variables shared between two rows
- try using distance matrix computed on the feature value with seriation
- add in feature selection methods to LIME input options
- think of a way to compute consistency across top two features
- Siggi suggests refitting the RF model to the perturbations and then continuing with LIME with the RF predictions from the new model - this may help to understand if the problems are due to the sampling procedure or LIME itself
- he also suggested looking into SMOTE for dealing the imbalance in the classes with sampling
- compare the simple models based on different number of bins using an F-test
- look into computing a diversity or consistency measure for the sensitivity analysis
- include a penalty for the number of parameters when choosing bins
- look at the AUC after binning
- compute a likelihood ratio prob TRUE / prob FALSE from the LIME ridge regression
- try visualizing the features from the test data using dimension reduction and
coloring them by variables suggested to be important by lime

**Explainer Models**

- read through papers giving an overview of the current explanation methods
- look into iml (https://www.youtube.com/watch?v=jP6Rg13PEkQ)
- look more into Molnar and his adviser Bernd
- look into [MAPLE](https://blog.ml.cmu.edu/2019/07/13/towards-interpretable-tree-ensembles/)

**Understanding LIME**

- edit details in understanding LIME journal
- look at the RF in the local region lime is considering to see if it agrees with the lime explanations
- try fitting LASSO logistic model and leave one out approach (for multicollinarity)
- assess lime results on logistic regression models
- try retriculate to apply python version of lime
- look into literature on binning methods
- think about why R^2 would be better for some binning methods
- read new paper on Anchor

**Possible Improvements to LIME**

- determine the best number of bins to use for each variable
- try out subsampling idea

**Simulation with RF**

- think about how to implement a simulation with a random forest
- look back at 601 notes from Kaiser for model assessment
- look at Hadley's 'removing the blindfold' for ideas
- read book Dr. Dixon lent me on sensitivity analyses

**Random Forests Confidence Intervals**

- look into papers by Giles Hooker and Lucas Mentch
- Dr. Nettleton paper

**Random Forest Interpretations**

- read https://www.r-bloggers.com/explaining-predictions-random-forest-post-hoc-analysis-randomforestexplainer-package/

## Repository Contents

Descriptions of the materials in this repository are listed below.

**code**

This folder is a place for storing code that is used throughout the project:

- `app.R`: code for the Shiny app for visualizing the LIME explanations associated with the bullet land comparisons
- `helper_functions.R`: R script with functions that I use in the journals and will probably end up using in the papers as well

**figures**

This folder contains static images:

- `lime.png`: fun picture of a lime to use in the readme
- `python-ridge1`, `python-ridge2`, and `python-ridge3`: screen shots of the python LIME package showing the use of a ridge regression as the explainer

**journals**

This folder contains folders with research journals.

- `00-objectives_and_ideas`: background of the research project, goals of the project, concerns with LIME, and ideas to try
- `01-hamby_data_and_models`: information on the Hamby data, cleaning of the training and testing data, visualizations of the data, and models fit to the data (rtrees and logistic regressions)
- `02-understanding_lime`: explanation of LIME procedure
- `03-applying_lime`: applications of LIME to the models fit to the Hamby data
- `04-assessing_lime`: visual diagnostics for assessing the lime explanations
- `05-iris_comparison`: applying LIME to a random forest fit to the iris data
- `06-literature_review`: notes on papers relating to LIME
- `07-computation_issues`: documentation of some of the computation issues I ran into during the project
- `08-logistic_regression`: examples of applying LIME to explain logistic regression models
- `09-refitting-rtrees`: examples where I retrain the rtrees model with the "current" version of the training data and compare the results

**main**

The documents contained in the top folder are as follows:

- `.gitattributes`: attributes relating to git
- `.gitignore`: files to ignore
- `notes from Heike.pdf`: scans of notes written by Heike for this project
- `README`: document containing to do list and contents of this repository