### 2020-09-25 journals/11-training-rtrees-with-CCFs_withlands/11-training-rtrees-with-CCFs_withlands.Rmd

---
title: "Experiments with rtrees using `CCFs_withlands` data"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE
)
```

This journal contains the results from 100 new random forests to the bullet training data observations where `flag == FALSE`.

```{r}
library(cowplot)
library(dplyr)
library(ggplot2)
library(purrr)
library(randomForest)
library(tidyr)
```

# Data

Load the data Heike extracted from the CSAFE database (on 9/23) which we currently believe is the correct training data for `rtrees`:

```{r}
CCFs_withlands <- read.csv("../../../diagnostics-paper/data/raw/CCFs_withlands.csv")
```

Load the random forest `rtrees` from the `bulletxtrctr` package:

```{r}
rtrees <- bulletxtrctr::rtrees
```

Extract the features used to fit `rtrees`:

```{r}
bullet_features <- rownames(bulletxtrctr::rtrees$importance)
bullet_features
```

# Random Forests

Randomly select a large amount of seeds (using a random seed):

```{r}
set.seed(20200925)
seeds = sample(x = 10000000:100000000, size = 250, replace = FALSE)
seeds
```

Function for fitting random forests to mimic rtrees with different seeds:

```{r}
fit_rf <- function(seed) {
  set.seed(seed)
  bullet_rf <-
    randomForest(
      y = factor(CCFs_withlands$match),
      x = CCFs_withlands %>% select(all_of(bullet_features)),
      ntree = bulletxtrctr::rtrees$ntree,
      mtry = bulletxtrctr::rtrees$mtry
    )  
}
```

Fit 100 new random forests and save or load:

```{r}
# Fit (and save) or load random forests
CCFs_withlands_rfs_file = "../../../data/CCFs_withlands_rfs.rds"
if (!file.exists(CCFs_withlands_rfs_file)) {
  CCFs_withlands_rfs = map(.x = seeds, .f = fit_rf)
  saveRDS(object = CCFs_withlands_rfs, file = CCFs_withlands_rfs_file)
} else {
  CCFs_withlands_rfs = readRDS(CCFs_withlands_rfs_file)
}
```

Function for extracting the confusion matrices from the model and put in a nice data frame:

```{r}
extract_confus <- function(model) {
  data.frame(model$confusion) %>%
  rename("FALSE" = "FALSE.", "TRUE" = "TRUE.", "class_error" = "class.error") %>%
  mutate(observed = rownames(model$confusion)) %>%
  pivot_wider(names_from = observed, values_from = everything())  
}
```

Function for extracting the importance values, computing the rank of the features based on importance, and putting in a data frame:

```{r}
extract_importance <- function(model) {
  data.frame(model$importance) %>%
    mutate(features = rownames(model$importance)) %>%
    arrange(desc(MeanDecreaseGini)) %>%
    mutate(rank = 1:n()) %>%
    rename("importance" = "MeanDecreaseGini") %>%
    select(features, importance, rank)
}
```

Extract the confusion matrices and importance values from all random forest models:

```{r}
res_confus <- map_df(.x = CCFs_withlands_rfs, .f = extract_confus, .id = "rf_model_id")
res_importance <- map_df(.x = CCFs_withlands_rfs, .f = extract_importance, .id = "rf_model_id")
```

# Visualizations of Results

## Class Errors

Obtain the class errors from `rtrees` and the paper random forest:

```{r}
comparison_values <-
  data.frame(
    model = "rtrees",
    class_error_true = rtrees$confusion[2, 3],
    class_error_false = rtrees$confusion[1, 3]
  )
```

Histograms of the class errors (true and false) show that the rtrees values appear to be in the extreme (false) or outside of the new random forest models (true):

```{r fig.width = 10, fig.height = 4.5}
hist_true <- 
  res_confus %>%
  ggplot(aes(x = class_error_TRUE)) +
  geom_histogram() +
  geom_vline(data = comparison_values,
             mapping = aes(xintercept = class_error_true, color = model))

hist_false <- 
  res_confus %>%
  ggplot(aes(x = class_error_FALSE)) +
  geom_histogram() +
  geom_vline(data = comparison_values,
             mapping = aes(xintercept = class_error_false, color = model))

plot_grid(hist_true, hist_false)
```

Class error for TRUE versus class error for FALSE and again the rtrees observations falls outside the range of the newly trained random forest models:

```{r}
ggplot() +
  geom_point(data = res_confus,
             mapping = aes(x = class_error_FALSE, y = class_error_TRUE)) +
  geom_point(
    data = comparison_values,
    mapping = aes(x = class_error_false, y = class_error_true, color = model)
  )
```

<!-- ## Variable Importance  -->

<!-- Prepare variable importance from `rtrees` and paper models for plotting: -->

<!-- ```{r} -->
<!-- vi_ranks <-  -->
<!--   vi %>% -->
<!--   arrange(desc(rtrees)) %>% -->
<!--   mutate(rtrees_rank = 1:n()) %>%  -->
<!--   arrange(desc(paper)) %>% -->
<!--   mutate(paper_rank = 1:n()) %>% -->
<!--   select(-rtrees, -paper) %>% -->
<!--   pivot_longer(names_to = "rf_model_id", values_to = "rank", cols = -features) -->
<!-- ``` -->

<!-- Tile plot showing the variable importance ranks for each model showing variability across newly trained models and some agreement (ex: rough_cor and ccf) and some disagreement (ex: sum_peaks and non_cms) with `rtrees`: -->

<!-- ```{r fig.width = 10, fig.height = 4.5} -->
<!-- rtrees_var_order <- -->
<!--   vi_ranks %>% -->
<!--   filter(rf_model_id == "rtrees_rank") %>% -->
<!--   arrange(desc(rank)) %>% -->
<!--   pull(features) -->

<!-- tile_org <- -->
<!--   vi_ranks %>% -->
<!--   mutate( -->
<!--     rank = factor(rank, levels = 1:9), -->
<!--     features = factor(features, levels = rtrees_var_order),  -->
<!--     rf_model_id = factor(rf_model_id, levels = c("rtrees_rank", "paper_rank")) -->
<!--   ) %>% -->
<!--   ggplot(aes(x = rf_model_id, y = features, fill = rank)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_brewer(palette = "RdYlBu") -->

<!-- tile_new <- -->
<!--   res_importance %>% -->
<!--   mutate( -->
<!--     rank = factor(rank, levels = 1:9), -->
<!--     features = factor(features, levels = rtrees_var_order) -->
<!--   ) %>% -->
<!--   ggplot(aes(x = rf_model_id, y = features, fill = rank)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_brewer(palette = "RdYlBu") -->

<!-- plot_grid(tile_org, tile_new) -->
<!-- ``` -->

<!-- Box plot of ranks across the models with colored points added for the `rtrees` and paper model variable importance ranks (features ordered by `rtrees` variable importance): -->

<!-- ```{r} -->
<!-- res_importance %>% -->
<!--   mutate(features = factor(features, levels = rtrees_var_order)) %>% -->
<!--   ggplot() + -->
<!--   geom_boxplot(aes(x = rank, y = features)) + -->
<!--   geom_jitter( -->
<!--     data = vi_ranks, -->
<!--     mapping = aes(x = rank, y = features, color = rf_model_id), -->
<!--     width = 0, -->
<!--     height = 0.2 -->
<!--   ) -->
<!-- ``` -->

<!-- Features versus variable importance ranks for new models (black lines and points) and `rtrees` and paper model (features ordered by `rtrees` variable importance): -->

<!-- ```{r} -->
<!-- res_importance %>% -->
<!--   mutate( -->
<!--     rank = factor(rank, levels = 1:9), -->
<!--     features = factor(features, levels = rtrees_var_order) -->
<!--   ) %>% -->
<!--   ggplot(aes(x = rank, y = features, group = rf_model_id)) + -->
<!--   geom_line(alpha = 0.2, size = 2) + -->
<!--   geom_point(alpha = 0.2, size = 3) + -->
<!--   geom_line( -->
<!--     data = vi_ranks, -->
<!--     mapping = aes( -->
<!--       x = rank, -->
<!--       y = features, -->
<!--       group = rf_model_id, -->
<!--       color = rf_model_id -->
<!--     ) -->
<!--   ) + -->
<!--   geom_jitter( -->
<!--     data = vi_ranks, -->
<!--     mapping = aes(x = rank, y = features, color = rf_model_id), -->
<!--     width = 0.1, -->
<!--     height = 0 -->
<!--   ) + -->
<!--   labs(title = "new random forest lines and points created with alpha shading") -->
<!-- ``` -->