### 2020-09-26 journals/09-identifying-rtrees-training-data/09-identifying-rtrees-training-data.Rmd
still a mess but did some more work

---
title: "Identifying `rtrees` Training Data"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE, 
  dpi = 300
)
```

This journal contains work to identify the data used to train `rtrees` (contained in both `bulletr` and `bulletxtrctr`). We recently discovered some discrepancies between `rtrees` and the training data we were using (`hamby-comparisons.csv`, previously `features-hamby173and252.csv`). Unfortunately, the original work by Eric is not well documented, so it is difficult to know for sure which dataset was used to train the model. However, Eric recently pointed us to some [code](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R), which he believes was used to train `rtrees`. Heike was able to rerun part of the code to extract data from the CSAFE database that was created by the first part of the code (scans to features). The new data (`CCFs_withlands.csv`) contains the same number of rows the number of predictions associated with `rtrees`. I have gathered all of our work trying to determine that `CCFs_withlands` is the data used to train `rtrees` in this journal.

```{r}
library(cowplot)
library(dplyr)
library(ggplot2)
library(purrr)
library(randomForest)
library(stringr)
library(tidyr)
```

# Raw Datasets

This section contains code for loading and performing initial comparisons of the raw datasets.

## Load Data

Load the data provided by Heike and CSAFE that we have been using and believed to be the training data for rtrees:

```{r}
hamby_comparisons_raw <- read.csv("../../../data/raw/hamby-comparisons.csv")
```

Load the data extracted (on 2020-09-23) from the CSAFE database and corresponds to the [R script](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R) Eric believes to be the one used to train `rtrees`.

```{r}
CCFs_withlands_raw <- read.csv("../../../data/raw/CCFs_withlands.csv")
```

Obtain the features used to train `rtrees` for use throughout this journal:

```{r}
rtrees_features <- rownames(bulletxtrctr::rtrees$importance)
```

## Initial Comparisons

Dimensions of the raw datasets:

```{r}
dim(hamby_comparisons_raw)
dim(CCFs_withlands_raw)
```

Note that `CCFs_withlands` has the same number of rows as the number of predictions associated with `rtrees`:

```{r}
length(bulletxtrctr::rtrees$predicted)
```

Comparing summaries of the distributions of the features used to train `rtrees` from the two datasets. Note that the main differences are seen with the distributions of `D` (distance) and `sd_D` (standard deviation of distance):

```{r}
summary(hamby_comparisons_raw %>% select(all_of(rtrees_features)))
summary(CCFs_withlands_raw %>% select(all_of(rtrees_features)))
```

This result is even true when the observations in `hamby_comparisons` known to have tank rash (`flag == FALSE`) are removed. This seems to suggest that the units of `D` and `sd_D` changed from `CCFs_withlands`:

```{r}
summary(hamby_comparisons_raw %>% filter(flag == FALSE) %>% select(all_of(rtrees_features)))
```

Comparing visualizations of the distributions of the features used to train `rtrees` from the two datasets to again see the differences in distributions of `D` and `sd_D`:

```{r fig.width = 10, fig.height = 3}
bind_rows(
  hamby_comparisons_raw %>%
    filter(flag == FALSE) %>%
    select(all_of(rtrees_features)) %>%
    pivot_longer(cols = everything()) %>%
    mutate(dataset = "hamby_comparisons"),
  CCFs_withlands_raw %>% select(all_of(rtrees_features)) %>%
    pivot_longer(cols = everything()) %>%
    mutate(dataset = "CCFs_withlands")
) %>%
  ggplot(aes(x = value, fill = dataset)) + 
  geom_histogram() +
  facet_grid(dataset ~ name)
```

# Data Cleaning

There are discrepancies in the naming conventions of the two datasets. This section contains the code that identifies the differences and cleans the data.

## Extracting Variables of Interest

The following variables are contained in the raw versions of the data. Note that they have different variables and different names. For our analysis, we only need the variables identifying the lands, the features used to train `rtrees`, the ground truth for whether or not the bullets are a match, and variables flagging bullets with tank rash.

```{r}
names(hamby_comparisons_raw)
names(CCFs_withlands_raw)
```

The code below separates the land id variables in `hamby_comparisons_raw` into separate columns for study, barrel, bullet, and land and selects only the variables of interest.

```{r}
hamby_comparisons_select <- 
  hamby_comparisons_raw %>%
  separate(land_id1, c("study1", "barrel1", "bullet1", "land1")) %>%
  separate(land_id2, c("study2", "barrel2", "bullet2", "land2")) %>%
  select(
    study1,
    barrel1,
    bullet1,
    land1,
    study2,
    barrel2,
    bullet2,
    land2,
    all_of(rtrees_features),
    same_source,
    flag
  )
```

The code below renames some of the variables in `CCFs_withlands_raw`, selects only the variables of interest, and converts all land labels to characters.

```{r}
CCFs_withlands_select <- 
  CCFs_withlands_raw %>%
  rename(
    "study1" = "study.x",
    "barrel1" = "barrel.x",
    "bullet1" = "bullet.x",
    "land1" = "land.x",
    "study2" = "study.y",
    "barrel2" = "barrel.y",
    "bullet2" = "bullet.y",
    "land2" = "land.y",
    "same_source"= "match"
  ) %>%
  select(
    study1,
    barrel1,
    bullet1,
    land1,
    study2,
    barrel2,
    bullet2,
    land2,
    all_of(rtrees_features),
    same_source
  ) %>%
  mutate(across(c(bullet1, land1, bullet2, land2), as.character))
```

Check to make sure the structures of the two datasets agree:

```{r}
hamby_comparisons_select %>% str()
CCFs_withlands_select %>% str()
```

## Matching Labels

The function below extracts the labels from the two datasets for a given variable name `var` and creates a plot to show the discrepancies between `hamby_comparison_select` and `CCFs_withlands_select`.

```{r}
plot_labels <- function(var, ccf_data) {

  # Extract the labels
  hc_labels <-
    unique(c(
      hamby_comparisons_select %>% pull(paste0(var, "1")),
      hamby_comparisons_select %>% pull(paste0(var, "2"))
    ))
  ccfwl_labels <-
    unique(c(
      ccf_data %>% pull(paste0(var, "1")),
      ccf_data %>% pull(paste0(var, "2"))
    ))

  # Plot the labels
  data.frame(data = c(
    rep("hamby comparisons", length(hc_labels)),
    rep("CCFs with lands", length(ccfwl_labels))
  ),
  labels = c(hc_labels, ccfwl_labels)) %>%
    ggplot(aes(x = labels, y = data)) +
    geom_tile() +
    labs(x = "", y = "", title = paste("Comparing labels in variable:", var))  
}
```

The plots below show that the following changes need to be made to `CCFs_withlands_select` in order to match with `hamby_comparison_select`: 
  - study label of `Hamby44` needs to be changed to `Hamby173`
  - lettered barrels need to be changed to `BrUnk` and `Br` needs to be added to other barrels
  - barrel letters need to be made the bullet label and `B` needs to be added to all bullets
  - `L` needs to be added to lands

```{r fig.width = 14, fig.height = 2.5}
plot_labels(var = "study", ccf_data = CCFs_withlands_select)
plot_labels(var = "barrel", ccf_data = CCFs_withlands_select)
plot_labels(var = "bullet", ccf_data = CCFs_withlands_select)
plot_labels(var = "land", ccf_data = CCFs_withlands_select)
```

Determine the letters in the `CCFs_withlands_select` unkown barrels:

```{r}
all_barrel_labels <-
  unique(c(
    as.character(CCFs_withlands_select$barrel1),
    as.character(CCFs_withlands_select$barrel2)
  ))
letters <- all_barrel_labels[!(all_barrel_labels %in% 1:10)]
letters
```

Clean `CCFs_withlands_select` so the labels match `bullet_train_raw`:

```{r}
CCFs_withlands_labelled <- 
  CCFs_withlands_select %>%
  mutate(
    study1 = ifelse(study1 == "Hamby44", "Hamby173", study1),
    study2 = ifelse(study2 == "Hamby44", "Hamby173", study2),
    bullet1 = ifelse(barrel1 %in% letters, as.character(barrel1), as.character(bullet1)),
    barrel1 = ifelse(barrel1 %in% letters, "Unk", barrel1),
    bullet2 = ifelse(barrel2 %in% letters, as.character(barrel2), as.character(bullet2)),
    barrel2 = ifelse(barrel2 %in% letters, "Unk", barrel2)
    ) %>%
  mutate(
    barrel1 = paste0("Br", barrel1),
    bullet1 = paste0("B", bullet1),
    land1 = paste0("L", land1),
    barrel2 = paste0("Br", barrel2),
    bullet2 = paste0("B", bullet2),
    land2 = paste0("L", land2)
  )
```

The plots below show that the labels are now in agreement:

```{r fig.width = 14, fig.height = 2.5}
plot_labels(var = "study", ccf_data = CCFs_withlands_labelled)
plot_labels(var = "barrel", ccf_data = CCFs_withlands_labelled)
plot_labels(var = "bullet", ccf_data = CCFs_withlands_labelled)
plot_labels(var = "land", ccf_data = CCFs_withlands_labelled)
```

## Final Touches

Create the land ids for both datasets:

```{r}
hamby_comparisons <-
  hamby_comparisons_select %>%
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  select(land_id1, land_id2, all_of(rtrees_features), same_source, flag)

CCFs_withlands <-
  CCFs_withlands_labelled %>%
  mutate(
    land_id1 = paste(study1, barrel1, bullet1, land1, sep = "-"),
    land_id2 = paste(study2, barrel2, bullet2, land2, sep = "-")
  ) %>%
  select(land_id1, land_id2, all_of(rtrees_features), same_source)
```

The structures of the cleaned data:

```{r}
hamby_comparisons %>% str()
CCFs_withlands %>% str()
```

# Land ID Differences

This section considers the differences in bullets included in the two datasets. The code below extracts the unique land IDs from the two datasets:

```{r}
hamby_comparisons_ids = unique(c(hamby_comparisons$land_id1, hamby_comparisons$land_id2))
CCFs_withlands_ids = unique(c(CCFs_withlands$land_id1, CCFs_withlands$land_id2))
```

Below are the number of bullets contained in each of the datasets. There are less in `CCFs_withlands` than `hamby_comparisons`:

```{r}
length(hamby_comparisons_ids)
length(CCFs_withlands_ids)
```

Identify the land ID in `CCFs_withlands` but not in `hamby-comparisons`:

```{r}
CCFs_withlands_ids[!(CCFs_withlands_ids %in% hamby_comparisons_ids)]
```

Identify the land ID in `hamby-comparisons` but not in `CCFs_withlands`:

```{r}
hamby_comparisons_ids[!(hamby_comparisons_ids %in% CCFs_withlands_ids)]
```

# Comparisons to rtrees

## Dimensions

One of the aspects which led to the realization that the `hamby-comparisons` data is not the training data for `rtrees` is that the number of observations in the data do not agree with the number of predictions in the `rtrees` model. The number of observations used to train `rtrees` is the following:

```{r}
length(bulletxtrctr::rtrees$predicted)
```

Here are the dimensions of the `hamby_comparisons` data which does not agree with the number of observations used to train `rtrees`:

```{r}
hamby_comparisons %>% dim()
```

Even with the observations removed that are known to have tank rash (`flag != FALSE`), the dimensions of the `hamby_comparisons` data do not agree with `rtrees`:

```{r}
hamby_comparisons %>% filter(flag == FALSE) %>% dim()
```

In the Hare, Hofmann, and Carriquiry (2017), they describe removing four land impressions that were flagged for quality assessment:

- Barrel 6 Bullet 2-1
- Barrel 9 Bullet 2-4
- Unknown Bullet B-2
- Unknown Bullet Q-4

These lands should correspond to Hamby 252. However, it is possible to obtain the same number of observations as used to train `rtrees` by filtering out observations in `hamby_comparisons` using these barrel, bullet, and land combinations without specifying the study (that is removing observations from both Hamby 173 and Hamby 252).

```{r}
hamby_comparisons_filtered <- 
  hamby_comparisons %>%
  mutate(bbl1 = str_remove(land_id1, pattern = "Hamby252-|Hamby173-"),
         bbl2 = str_remove(land_id2, pattern = "Hamby252-|Hamby173-")) %>%
  filter(!(bbl1 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4") | 
           bbl2 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4"))) %>%
  select(-bbl1, -bbl2)
hamby_comparisons_filtered %>% dim()
```

While it is possible to use `hamby_comparisons` to get to the same number of observations as `rtrees`, it is only a guess as to how the bullets were removed. On the other hand, the number of rows in `CCFs_withlands` already has the same number of rows as `rtrees`:

```{r}
CCFs_withlands %>% dim()
```

# Old Code

<!-- The code below removes observations based on Eric Hare's code (available [here](https://github.com/erichare/imaging-paper/blob/master/code/full_run.R)), which he thinks may be the code used to train `rtrees`: -->

<!-- ```{r} -->
<!-- bullet_train_eric_filter <-  -->
<!--   bullet_train_raw %>% -->
<!--   separate(col = "land_id1", into = c("study1", "barrel1", "bullet1", "land1")) %>% -->
<!--   separate(col = "land_id2", into = c("study2", "barrel2", "bullet2", "land2")) %>% -->
<!--   filter(study1 != "Hamby173" | bullet1 != "BE") %>% -->
<!--   filter(study2 != "Hamby173" | bullet2 != "BE") -->

<!-- # Eric's code for reference: -->
<!-- # filter(study.y != "Hamby44" | barrel.y != "E") %>% -->
<!-- # filter(study.x != "Hamby44" | barrel.x != "E") %>% -->
<!-- ``` -->

<!-- Here, the dimensions to not agree with the number of observations in `rtrees`: -->

<!-- ```{r} -->
<!-- dim(bullet_train_eric_filter) -->
<!-- length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_eric_filter)[1] -->
<!-- ``` -->

<!-- ```{r} -->
<!-- bullet_train_eric_filter_v2 <- -->
<!--   bullet_train_raw %>% -->
<!--   mutate(bbl1 = str_remove(land_id1, pattern = "Hamby252-|Hamby173-"), -->
<!--          bbl2 = str_remove(land_id2, pattern = "Hamby252-|Hamby173-")) %>% -->
<!--   filter(!(bbl1 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4") |  -->
<!--            bbl2 %in% c("Br6-B2-L1", "Br9-B2-L4", "BrUnk-BB-L2", "BrUnk-BQ-L4"))) %>% -->
<!--   select(-bbl1, -bbl2) %>% -->
<!--   separate(col = "land_id1", -->
<!--            into = c("study1", "barrel1", "bullet1", "land1")) %>% -->
<!--   separate(col = "land_id2", -->
<!--            into = c("study2", "barrel2", "bullet2", "land2")) %>% -->
<!--   filter(study1 != "Hamby173" | bullet1 != "BE") %>% -->
<!--   filter(study2 != "Hamby173" | bullet2 != "BE") -->

<!-- dim(bullet_train_eric_filter_v2) -->
<!-- length(bulletxtrctr::rtrees$predicted) == dim(bullet_train_eric_filter_v2)[1] -->
<!-- ``` -->

<!-- Extracting the predictions on the training data from `rtrees` and using `rtrees` to make predictions on the filtered data: -->

<!-- ```{r} -->
<!-- rtrees_pred = predict(bulletxtrctr::rtrees, type = "prob")[,2] -->

<!-- new_pred = predict(bulletxtrctr::rtrees,  -->
<!--                    bullet_train_filtered %>%  -->
<!--                      select(rownames(bulletxtrctr::rtrees$importance)), type = "prob")[,2] -->

<!-- new_pred_full = predict(bulletxtrctr::rtrees,  -->
<!--                         bullet_train_raw %>% -->
<!--                           select(rownames(bulletxtrctr::rtrees$importance)), type = "prob")[,2] -->
<!-- ``` -->

<!-- Plot comparing the two predictions (we discovered that the `predict` function from random forest returns different results when applied to only the model (out-of-bag predictions) and the training data (regular predictions), so this is not a good way to test for agreement between training data and models): -->

<!-- ```{r} -->
<!-- data.frame(rtrees_pred = sort(rtrees_pred), new_pred = sort(new_pred)) %>% -->
<!--   ggplot(aes(rtrees_pred, new_pred)) +  -->
<!--   geom_point() -->
<!-- ``` -->

<!-- Plot showing distributions of predictions from `rtrees` and new predictions made on full training data (same comment here as above about not being able to use this to determine if we have the correct training data for the model): -->

<!-- ```{r} -->
<!-- data.frame(data = c(rep("rtrees", length(rtrees_pred)),  -->
<!--                     rep("new predictions on full data", length(new_pred_full))), pred = c(rtrees_pred, new_pred_full)) %>% -->
<!--   ggplot(aes(x = pred, fill = data)) +  -->
<!--   geom_density(alpha = 0.75) -->
<!-- ``` -->

<!-- Comparison to the data Eric Hare said was used to fit `rtrees` which has half the number of observations as the number of rtrees predictions:  -->

<!-- ```{r} -->
<!-- eric <- read.csv("https://raw.githubusercontent.com/erichare/imaging-paper/master/data/data-25-25/bullet-stats.csv") -->
<!-- dim(eric) -->
<!-- ``` -->

<!-- New data plot: -->

<!-- ```{r} -->
<!-- preds <- -->
<!--   data.frame( -->
<!--     oob = predict(bulletxtrctr::rtrees, type = "prob")[, 2], -->
<!--     all_data = predict( -->
<!--       bulletxtrctr::rtrees, -->
<!--       ccfs_withlands_raw %>% select(all_of(rf_features)), -->
<!--       type = "prob" -->
<!--     )[, 2] -->
<!--   ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- preds %>%  -->
<!--   ggplot(aes(x = all_data, y = oob)) +  -->
<!--   geom_point() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- preds %>% -->
<!--   mutate(obs = 1:n()) %>% -->
<!--   pivot_longer(cols = -obs) %>% -->
<!--   ggplot(aes(x = value, fill = name)) +  -->
<!--   geom_density(alpha = 0.5) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- preds %>%  -->
<!--   mutate(diff = oob - all_data) %>% -->
<!--   ggplot(aes(x = diff)) +  -->
<!--   geom_histogram() -->
<!-- ``` -->
<!-- # Data -->

<!-- Load the prepared training data which only has observations with `flag == FALSE` (prepared in the LIME diagnostics paper repository): -->

<!-- ```{r} -->
<!-- bullet_train <- read.csv("../../../diagnostics-paper/data/bullet-train.csv") -->
<!-- ``` -->

<!-- Load the random forest `rtrees` from the `bulletxtrctr` package and the random forest model I recently fit using the `bullet_train` data in the paper: -->

<!-- ```{r} -->
<!-- rtrees <- bulletxtrctr::rtrees -->
<!-- paper_rf <- readRDS("../../../diagnostics-paper/data/bullet-rf.rds") -->
<!-- ``` -->

<!-- Extract the features used to fit `rtrees`: -->

<!-- ```{r} -->
<!-- bullet_features <- rownames(bulletxtrctr::rtrees$importance) -->
<!-- bullet_features -->
<!-- ``` -->

<!-- # `rtrees` versus paper model -->

<!-- The code below was used to fit the random forest in the paper but is not run here: -->

<!-- ```{r eval = FALSE} -->
<!-- set.seed(20160104) -->
<!-- bullet_rf <- -->
<!--   randomForest( -->
<!--     y = factor(bullet_train$samesource), -->
<!--     x = bullet_train %>% select(all_of(bullet_features)), -->
<!--     ntree = bulletxtrctr::rtrees$ntree, -->
<!--     mtry = bulletxtrctr::rtrees$mtry -->
<!--   ) -->
<!-- ``` -->

<!-- Confusion matrices from the two models: -->

<!-- ```{r} -->
<!-- rtrees$confusion -->
<!-- paper_rf$confusion -->
<!-- ``` -->

<!-- Variable importance values: -->

<!-- ```{r} -->
<!-- vi = data.frame( -->
<!--   features = rownames(rtrees$importance), -->
<!--   rtrees = rtrees$importance[, 1], -->
<!--   paper = paper_rf$importance[, 1] -->
<!-- ) -->
<!-- ``` -->

<!-- # New Random Forests -->

<!-- Randomly select 100 seeds (using a random seed):  -->

<!-- ```{r} -->
<!-- set.seed(20200917) -->
<!-- seeds = sample(x = 10000000:100000000, size = 100, replace = FALSE) -->
<!-- seeds -->
<!-- ``` -->

<!-- Function for fitting random forests to mimic rtrees with different seeds: -->

<!-- ```{r} -->
<!-- fit_rf <- function(seed) { -->
<!--   set.seed(seed) -->
<!--   bullet_rf <- -->
<!--     randomForest( -->
<!--       y = factor(bullet_train$samesource), -->
<!--       x = bullet_train %>% select(all_of(bullet_features)), -->
<!--       ntree = bulletxtrctr::rtrees$ntree, -->
<!--       mtry = bulletxtrctr::rtrees$mtry -->
<!--     )   -->
<!-- } -->
<!-- ``` -->

<!-- Fit 100 new random forests and save or load: -->

<!-- ```{r} -->
<!-- # Fit (and save) or load random forests -->
<!-- bullet_rfs_file = "../../../data/bullet_rfs.rds" -->
<!-- if (!file.exists(bullet_rfs_file)) { -->
<!--   bullet_rfs = map(.x = seeds, .f = fit_rf) -->
<!--   saveRDS(object = bullet_rfs, file = bullet_rfs_file) -->
<!-- } else { -->
<!--   bullet_rfs = readRDS(bullet_rfs_file) -->
<!-- } -->
<!-- ``` -->

<!-- Function for extracting the confusion matrices from the model and put in a nice data frame: -->

<!-- ```{r} -->
<!-- extract_confus <- function(model) { -->
<!--   data.frame(model$confusion) %>% -->
<!--   rename("FALSE" = "FALSE.", "TRUE" = "TRUE.", "class_error" = "class.error") %>% -->
<!--   mutate(observed = rownames(model$confusion)) %>% -->
<!--   pivot_wider(names_from = observed, values_from = everything())   -->
<!-- } -->
<!-- ``` -->

<!-- Function for extracting the importance values, computing the rank of the features based on importance, and putting in a data frame:  -->

<!-- ```{r} -->
<!-- extract_importance <- function(model) { -->
<!--   data.frame(model$importance) %>% -->
<!--     mutate(features = rownames(model$importance)) %>% -->
<!--     arrange(desc(MeanDecreaseGini)) %>% -->
<!--     mutate(rank = 1:n()) %>% -->
<!--     rename("importance" = "MeanDecreaseGini") %>% -->
<!--     select(features, importance, rank) -->
<!-- } -->
<!-- ``` -->

<!-- Extract the confusion matrices and importance values from all random forest models: -->

<!-- ```{r} -->
<!-- res_confus <- map_df(.x = bullet_rfs, .f = extract_confus, .id = "rf_model_id") -->
<!-- res_importance <- map_df(.x = bullet_rfs, .f = extract_importance, .id = "rf_model_id") -->
<!-- ``` -->

<!-- # Visualizations of Results -->

<!-- ## Class Errors -->

<!-- Obtain the class errors from `rtrees` and the paper random forest: -->

<!-- ```{r} -->
<!-- comparison_values <- -->
<!--   data.frame( -->
<!--     model = c("paper", "rtrees"), -->
<!--     class_error_true = c(paper_rf$confusion[2, 3], rtrees$confusion[2, 3]), -->
<!--     class_error_false = c(paper_rf$confusion[1, 3], rtrees$confusion[1, 3]) -->
<!--   ) -->
<!-- ``` -->

<!-- Histograms of the class errors (true and false) show that the rtrees values appear to be in the extreme (false) or outside of the new random forest models (true): -->

<!-- ```{r fig.width = 10, fig.height = 4.5} -->
<!-- hist_true <-  -->
<!--   res_confus %>% -->
<!--   ggplot(aes(x = class_error_TRUE)) + -->
<!--   geom_histogram() + -->
<!--   geom_vline(data = comparison_values, -->
<!--              mapping = aes(xintercept = class_error_true, color = model)) -->

<!-- hist_false <-  -->
<!--   res_confus %>% -->
<!--   ggplot(aes(x = class_error_FALSE)) + -->
<!--   geom_histogram() + -->
<!--   geom_vline(data = comparison_values, -->
<!--              mapping = aes(xintercept = class_error_false, color = model)) -->

<!-- plot_grid(hist_true, hist_false) -->
<!-- ``` -->

<!-- Class error for TRUE versus class error for FALSE and again the rtrees observations falls outside the range of the newly trained random forest models: -->

<!-- ```{r} -->
<!-- ggplot() + -->
<!--   geom_point(data = res_confus, -->
<!--              mapping = aes(x = class_error_FALSE, y = class_error_TRUE)) + -->
<!--   geom_point( -->
<!--     data = comparison_values, -->
<!--     mapping = aes(x = class_error_false, y = class_error_true, color = model) -->
<!--   ) -->
<!-- ``` -->

<!-- ## Variable Importance  -->

<!-- Prepare variable importance from `rtrees` and paper models for plotting: -->

<!-- ```{r} -->
<!-- vi_ranks <-  -->
<!--   vi %>% -->
<!--   arrange(desc(rtrees)) %>% -->
<!--   mutate(rtrees_rank = 1:n()) %>%  -->
<!--   arrange(desc(paper)) %>% -->
<!--   mutate(paper_rank = 1:n()) %>% -->
<!--   select(-rtrees, -paper) %>% -->
<!--   pivot_longer(names_to = "rf_model_id", values_to = "rank", cols = -features) -->
<!-- ``` -->

<!-- Tile plot showing the variable importance ranks for each model showing variability across newly trained models and some agreement (ex: rough_cor and ccf) and some disagreement (ex: sum_peaks and non_cms) with `rtrees`: -->

<!-- ```{r fig.width = 10, fig.height = 4.5} -->
<!-- rtrees_var_order <- -->
<!--   vi_ranks %>% -->
<!--   filter(rf_model_id == "rtrees_rank") %>% -->
<!--   arrange(desc(rank)) %>% -->
<!--   pull(features) -->

<!-- tile_org <- -->
<!--   vi_ranks %>% -->
<!--   mutate( -->
<!--     rank = factor(rank, levels = 1:9), -->
<!--     features = factor(features, levels = rtrees_var_order),  -->
<!--     rf_model_id = factor(rf_model_id, levels = c("rtrees_rank", "paper_rank")) -->
<!--   ) %>% -->
<!--   ggplot(aes(x = rf_model_id, y = features, fill = rank)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_brewer(palette = "RdYlBu") -->

<!-- tile_new <- -->
<!--   res_importance %>% -->
<!--   mutate( -->
<!--     rank = factor(rank, levels = 1:9), -->
<!--     features = factor(features, levels = rtrees_var_order) -->
<!--   ) %>% -->
<!--   ggplot(aes(x = rf_model_id, y = features, fill = rank)) + -->
<!--   geom_tile() + -->
<!--   scale_fill_brewer(palette = "RdYlBu") -->

<!-- plot_grid(tile_org, tile_new) -->
<!-- ``` -->

<!-- Box plot of ranks across the models with colored points added for the `rtrees` and paper model variable importance ranks (features ordered by `rtrees` variable importance): -->

<!-- ```{r} -->
<!-- res_importance %>% -->
<!--   mutate(features = factor(features, levels = rtrees_var_order)) %>% -->
<!--   ggplot() + -->
<!--   geom_boxplot(aes(x = rank, y = features)) + -->
<!--   geom_jitter( -->
<!--     data = vi_ranks, -->
<!--     mapping = aes(x = rank, y = features, color = rf_model_id), -->
<!--     width = 0, -->
<!--     height = 0.2 -->
<!--   ) -->
<!-- ``` -->

<!-- Features versus variable importance ranks for new models (black lines and points) and `rtrees` and paper model (features ordered by `rtrees` variable importance): -->

<!-- ```{r} -->
<!-- res_importance %>% -->
<!--   mutate( -->
<!--     rank = factor(rank, levels = 1:9), -->
<!--     features = factor(features, levels = rtrees_var_order) -->
<!--   ) %>% -->
<!--   ggplot(aes(x = rank, y = features, group = rf_model_id)) + -->
<!--   geom_line(alpha = 0.2, size = 2) + -->
<!--   geom_point(alpha = 0.2, size = 3) + -->
<!--   geom_line( -->
<!--     data = vi_ranks, -->
<!--     mapping = aes( -->
<!--       x = rank, -->
<!--       y = features, -->
<!--       group = rf_model_id, -->
<!--       color = rf_model_id -->
<!--     ) -->
<!--   ) + -->
<!--   geom_jitter( -->
<!--     data = vi_ranks, -->
<!--     mapping = aes(x = rank, y = features, color = rf_model_id), -->
<!--     width = 0.1, -->
<!--     height = 0 -->
<!--   ) + -->
<!--   labs(title = "new random forest lines and points created with alpha shading") -->
<!-- ``` -->

# Session Info

```{r}
sessionInfo()
```