### 2020-09-29 README.Rmd
updating readme

---
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LIME Research <img align="right" height="200" src="./figures/lime.png">

Repository for research journals for my research projects related to LIME. See below for descriptions of the contents of this repository. For easy viewing of the journals, use these links:

0. [Objectives and Ideas]()
1. [Information on Hamby Data and Models]()
2. [Understanding LIME]()
3. [Applying LIME to Hamby Data]()
4. [Assessing LIME on Hamby Data]()
5. [Applying and Assessing LIME on Iris Data]()
6. [Notes on Readings]()
7. [Computational Issues]()
8. [LIME Applied to Logistic Regression]()
9. [Identifying `rtrees` Training Data]()
10. [LIME Variability Across Hamby Data and Models]()

## Repository Contents

Descriptions of the materials in this repository are listed below.

**code**

This folder is a place for storing code that is used throughout the project:

- `app.R`: code for the Shiny app for visualizing the LIME explanations associated with the bullet land comparisons
- `helper_functions.R`: R script with functions that I use in the journals and will probably end up using in the papers as well

**figures**

This folder contains static images:

- `lime.png`: fun picture of a lime to use in the readme
- `python-ridge1`, `python-ridge2`, and `python-ridge3`: screen shots of the python LIME package showing the use of a ridge regression as the explainer

**journals**

This folder contains folders with research journals.

- `00-objectives_and_ideas`: background of the research project, goals of the project, concerns with LIME, and ideas to try
- `01-hamby_data_and_models`: information on the Hamby data, cleaning of the training and testing data, visualizations of the data, and models fit to the data (rtrees and logistic regressions)
- `02-understanding_lime`: explanation of LIME procedure
- `03-applying_lime`: applications of LIME to the models fit to the Hamby data
- `04-assessing_lime`: visual diagnostics for assessing the lime explanations
- `05-iris_comparison`: applying LIME to a random forest fit to the iris data
- `06-literature_review`: notes on papers relating to LIME
- `07-computation_issues`: documentation of some of the computation issues I ran into during the project
- `08-logistic_regression`: examples of applying LIME to explain logistic regression models
- `09-refitting-rtrees`: examples where I retrain the rtrees model with the "current" version of the training data and compare the results

**main**

The documents contained in the top folder are as follows:

- `.gitattributes`: attributes relating to git
- `.gitignore`: files to ignore
- `notes from Heike.pdf`: scans of notes written by Heike for this project
- `README`: document containing to do list and contents of this repository