### 2020-09-29 old-journals/bullet data/journals/Heike_first example of lime with bullet data.Rmd

---
title: "Testing Lime"
author: "Heike Hofmann"
date: "10/7/2017"
output: pdf_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(knitr)
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  error = FALSE, 
  collapse = TRUE,
  comment = "",
  fig.height = 6,
  fig.width = 6,
  fig.align = "center",
  out.width= '\\textwidth',
  cache = FALSE,
  fig.path='figures-lime/',
  echo=TRUE
)
options(knitr.table.format = "latex")

library(tidyverse)
library(scales)
library(multidplyr)
library(bulletr)
library(gridExtra)
library(kableExtra)
```

```{r read}
hamby <- read.csv("data/features-hamby.csv.gz") %>% filter(study1 != "Cary", study2 != "Cary")
library(caret)
library(lime)

# Split up the data set
idx <- createDataPartition(hamby$match, p = 0.1, list=FALSE)

hamby_test <- hamby[idx,]
hamby_train <- hamby[-idx,]

# Create Random Forest model on hamby data
vars <- c("ccf", "rough_cor", "lag", "D", "sd_D", "overlap", "cms", "non_cms", "sum_peaks")
model <- train(hamby_train %>% select(one_of(vars)), factor(hamby_train$match), method = 'rf')
saveRDS(model, file="lime-data/model.rds")

# Create an explainer object
explainer <- lime(hamby_train %>% select(one_of(vars)), model)
saveRDS(explainer, file="lime-data/explainer.rds")

# Explain new observation
explanation <- explain(hamby_test %>% select(one_of(vars)), explainer, n_labels = 1, n_features = 9)
explanation$truth <- rep(hamby_test$match,each=9)
plot_features(explanation[1:9,])
saveRDS(explanation, file="lime-data/explanation.rds")
```

```{r second}
newGrid = expand.grid(mtry = 4:6)
second <- train(hamby_train %>% select(one_of(vars)), factor(hamby_train$match), method = 'rf', tuneGrid = newGrid)
saveRDS(second, file="lime-data/model-second.rds") # mtry=4 seems to work the best (by a very small margin, but there's not that much room for improvement anyway)
```