### 2020-09-29 journals/10-lime_hamby_variability/10-lime_hamby_variability.Rmd
added journal for comparing lime explanations across models

---
title: "LIME Variability Across Hamby Data and Model"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE
)
```

We have finally identified the data that was most likely used to train `rtrees` (`CCFs_withlands`). This data is different from the one we previously believed to be the one that trained `rtrees` (`hamby_comparisons`; previously called `features-hamby173and252.csv`). It is of interest to keep a record of how the LIME explanations on a testing dataset (Hamby 224 sets 1 and 11) vary across the various training datasets and models. This journal applies LIME to various versions of a random forest trained on the Hamby bullet signature comparisons data.

```{r}
library(dplyr)
library(forcats)
library(ggplot2)
library(gretchenalbrecht)
library(limeaid)
```

# Data

Load the filtered version of the `hamby_comparisons` data: 
  - LEAs with tank rash in Eric's paper filtered out
  - Contains same number of rows as `rtrees`
  - Prepared in journal [09-identifying-rtrees-training-data.Rmd](https://goodekat.github.io/LIME-research-journals/journals/09-identifying-rtrees-training-data/09-identifying-rtrees-training-data.html)

```{r}
hamby_comparisons_filtered <- read.csv("../../../data/hamby_comparisons_filtered.csv")
```

Load the cleaned version of the `CCFs_withlands` data (prepared in [09-identifying-rtrees-training-data.Rmd](https://goodekat.github.io/LIME-research-journals/journals/09-identifying-rtrees-training-data/09-identifying-rtrees-training-data.html)):

```{r}
CCFs_withlands_clean <- read.csv("../../../data/CCFs_withlands_clean.csv")
```

Load the bullet testing data using in the LIME diagnostics paper (prepared in https://github.com/goodekat/LIME-diagnostics-paper/blob/master/code/01-raw-file-compression.Rmd):

```{r}
bullet_test <- read.csv("../../../diagnostics-paper/data/bullet-test.csv")
```

Obtain the features used to train `rtrees` for use throughout this journal:

```{r}
rtrees_features <- rownames(bulletxtrctr::rtrees$importance)
```

# Apply LIME

## rtrees with hamby_comparisons

Wrong training data for `rtrees`

```{r}
# Create a file path
file_rtrees_hc_explanations = "../../../data/rtrees_explanations_hc.rds"

# Implement and save or load LIME explanations
if (!file.exists(file_rtrees_hc_explanations)) {

  # Apply LIME
  rtrees_hc_explanations <- apply_lime(
    train = hamby_comparisons_filtered %>% select(all_of(rtrees_features)),
    test = bullet_test %>% select(all_of(rtrees_features)),
    model = bulletxtrctr::rtrees,
    label = as.character(TRUE),
    n_features = 3,
    sim_method = c('quantile_bins', 'equal_bins',
                   'kernel_density', 'normal_approx'),
    nbins = 2:6,
    feature_select = "auto",
    dist_fun = "gower",
    kernel_width = NULL,
    gower_pow = c(0.5, 1, 10),
    return_perms = FALSE,
    all_fs = FALSE,
    seed = 20190914)

  # Save the explanations
  saveRDS(object = rtrees_hc_explanations, file = file_rtrees_hc_explanations)

} else {

  # Load the explanations
  rtrees_hc_explanations <- readRDS(file_rtrees_hc_explanations)

}
```

## rtrees with CCFs_withlands

Probably the training data for `rtrees`

```{r}
# Create a file path
file_rtrees_ccfwl_explanations = "../../../data/rtrees_explanations_ccfwl.rds"

# Implement and save or load LIME explanations
if (!file.exists(file_rtrees_ccfwl_explanations)) {

  # Apply LIME
  rtrees_ccfwl_explanations <- apply_lime(
    train = CCFs_withlands_clean %>% select(all_of(rtrees_features)),
    test = bullet_test %>% select(all_of(rtrees_features)),
    model = bulletxtrctr::rtrees,
    label = as.character(TRUE),
    n_features = 3,
    sim_method = c('quantile_bins', 'equal_bins',
                   'kernel_density', 'normal_approx'),
    nbins = 2:6,
    feature_select = "auto",
    dist_fun = "gower",
    kernel_width = NULL,
    gower_pow = c(0.5, 1, 10),
    return_perms = FALSE,
    all_fs = FALSE,
    seed = 20190914)

  # Save the explanations
  saveRDS(object = rtrees_ccfwl_explanations, file = file_rtrees_ccfwl_explanations)

} else {

  # Load the explanations
  rtrees_ccfwl_explanations <- readRDS(file_rtrees_ccfwl_explanations)

}
```

## New RF with CCFs_withlands

Using the data that was probably used to train `rtrees` to train a new random forest (see the [Rnw of the LIME diagnostics paper](https://github.com/goodekat/LIME-diagnostics-paper/blob/master/paper.Rnw) for the code):

```{r}
bullet_explain <- readRDS("../../../diagnostics-paper/data/bullet-explain.rds")
```

# Feature Heatmaps

## rtrees with hamby_comparisons

```{r fig.width = 12, fig.height = 12}
# Create a feature heatmap
plot_feature_heatmap(
  rtrees_hc_explanations$explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")
```

## rtrees with CCFs_withlands

```{r fig.width = 12, fig.height = 12}
# Create a feature heatmap
plot_feature_heatmap(
  rtrees_ccfwl_explanations$explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")
```

## New RF with CCFs_withlands

```{r fig.width = 12, fig.height = 12}
# Create a feature heatmap
plot_feature_heatmap(
  bullet_explain %>%
    mutate(
      label = as.factor(label),
      feature = fct_recode(
        feature,
        "Rough Correlation" = "rough_cor",
        "Consecutively Matching Striae" = "cms",
        "Distance" = "D",
        "Matches" = "matches",
        "Mismatches" = "mismatches",
        "Non-Consecutively Matching Striae" = "non_cms",
        "Cross Correlation Function" = "ccf",
        "Sum of Peaks" = "sum_peaks",
        "Distance Standard Deviation" = "sd_D"
      )
    ),
  facet_var = bullet_test %>%
    mutate(samesource = fct_recode(
      factor(samesource),
      "Match" = "TRUE",
      "Non-Match" = "FALSE"
    )) %>%
    pull(samesource),
  order_method = "PCA"
) +
  scale_fill_gretchenalbrecht(palette = "last_rays", discrete = TRUE) +
  scale_color_gretchenalbrecht(palette = "last_rays",
                               discrete = TRUE,
                               reverse = TRUE) +
  theme_bw() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    strip.background = element_rect(color = "white", fill = "white"),
    strip.text.y.right = element_text(angle = 0),
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  guides(fill = guide_legend(nrow = 3)) +
  labs(y = "Case", color = "Complex Model Feature", fill = "Complex Model Feature")
```

# Assessment Metric Plots

## rtrees with hamby_comparisons

```{r}
plot_metrics(
  rtrees_hc_explanations$explain %>% mutate(label = as.factor(label)),
  add_lines = TRUE,
  line_alpha = 0.75
) +
  theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.placement = "outside"
  )
```

## rtrees with CCFs_withlands

```{r}
plot_metrics(
  rtrees_ccfwl_explanations$explain %>% mutate(label = as.factor(label)),
  add_lines = TRUE,
  line_alpha = 0.75
) +
  theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.placement = "outside"
  )
```

## New RF with CCFs_withlands

```{r}
plot_metrics(
  bullet_explain %>% mutate(label = as.factor(label)),
  add_lines = TRUE,
  line_alpha = 0.75
) +
  theme_bw() +
  theme(
    strip.background = element_rect(color = "white", fill = "white"),
    strip.placement = "outside"
  )
```

# Session Info

```{r}
sessionInfo()
```