### 2020-09-30 README.Rmd
small updates

---
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# LIME Research <img align="right" height="200" src="https://i.pinimg.com/originals/de/18/33/de18338e3313edd97d8156d987244e74.jpg">

Repository for research journals for my research projects related to LIME. See below for descriptions of the contents of this repository. For easy viewing of the journals, use these links:

0. [Objectives and Ideas](https://goodekat.github.io/LIME-research-journals/journals/00-objectives_and_ideas/00-objectives_and_ideas.html)
1. [Information on Hamby Data and Models](https://goodekat.github.io/LIME-research-journals/journals/01-hamby_data_and_models/01-hamby_data_and_models.html)
2. [Understanding LIME](https://goodekat.github.io/LIME-research-journals/journals/02-understanding_lime/02-understanding_lime.html)
3. [Applying LIME to Hamby Data](https://goodekat.github.io/LIME-research-journals/journals/03-applying_lime/03-applying_lime.html)
4. [Assessing LIME on Hamby Data](https://goodekat.github.io/LIME-research-journals/journals/04-assessing_lime/04-assessing_lime.html)
5. [Applying and Assessing LIME on Iris Data](https://goodekat.github.io/LIME-research-journals/journals/05-iris_comparison/05-iris_comparison.html)
6. [Notes on Readings](https://goodekat.github.io/LIME-research-journals/journals/06-literature_review/06-literature_review.html)
7. [Computational Issues](https://goodekat.github.io/LIME-research-journals/journals/07-computation_issues/07-computation_issues.html)
8. [LIME Applied to Logistic Regressions on Sine Data](https://goodekat.github.io/LIME-research-journals/journals/08-logistic_regression/08-logistic_regression.html)
9. [Identifying rtrees Training Data](https://goodekat.github.io/LIME-research-journals/journals/09-identifying_rtrees_training_data/09-identifying_rtrees_training_data.html)
10. [LIME Variability Across Hamby Data and Models](https://goodekat.github.io/LIME-research-journals/journals/10-lime_hamby_variability/10-lime_hamby_variability.html)

## Repository Contents

Descriptions of the materials in this repository are listed below.

[**code**](code/)

Place for storing R code used in the research journals:

- [app.R](code/app.R): code for the Shiny app for visualizing the LIME explanations associated with the bullet land comparisons
- [helper_functions.R](code/helper_functions.R): R script with functions that I use in the journals

[**figures**](figures)

Contains static images:

- [laugel-lime-concept.png](figures/laugel-lime-concept.png) and [laugel-locality-metric](figures/laugel-locality-metric.png): figures from the Laugel et al. (2018) paper
- [python-ridge1](figures/python-ridge1.png), [python-ridge2](figures/python-ridge2.png), and [python-ridge3](figures/python-ridge3.png): screen shots of the python LIME package showing the use of a ridge regression as the explainer

[**journals**](journals)

This folder contains folders with research journals.

- [00-objectives_and_ideas](journals/00-objectives_and_ideas): background of the research project, goals of the project, concerns with LIME, and ideas to try
- [01-hamby_data_and_models](journals/01-hamby_data_and_models)***: information on the Hamby data, cleaning of the training and testing data, visualizations of the data, and models fit to the data (rtrees and logistic regressions)
- [02-understanding_lime](journals/02-understanding_lime)***: work done to understand the LIME algorithm
- [03-applying_lime](journals/03-applying_lime)***: applications of LIME to the models fit to the Hamby data
- [04-assessing_lime](journals/04-assessing_lime)***: visual diagnostics for assessing the lime explanations
- [05-iris_comparison](journals/05-iris_comparison): applying LIME to a random forest fit to the iris data and assessing the explanations
- [06-literature_review](journals/06-literature_review/): notes on papers relating to LIME
- [07-computation_issues](journals/07-computation_issues/)***: documentation of some of the computation issues I ran into during the project
- [08-logistic_regression](journals/08-logistic_regression): examples of applying and diagnosing LIME with logistic regression models fit to the sine data
- [09-identifying_rtrees_training_data](journals/09-identifying_rtrees_training_data/): examples where I retrain the rtrees model with the "current" version of the training data and compare the results
- [10-lime_hamby_variability](journals/10-lime_hamby_variability/): investigating variability between LIME explanations using diagnostic figures for random forest models trained on the Hamby bullet data

*** **Note**: We have realized that the training dataset for the `rtrees` model is not the one used in these files. See [09-identifying_rtrees_training_data](journals/09-identifying_rtrees_training_data) for more details on the correct dataset. The correct data is used in the LIME diagnostics paper. For time reasons, the results in these journals have not been updated.

[**old-journals**](old-journals)

Contain materials from my earliest research journals relating to the LIME project. It is likely that the code in these files will not run. However, the knit versions of the R markdown documents may still be useful.

**main**

The documents contained in the top folder are as follows:

- [.gitignore](.gitignore): files to ignore
- [notes from Heike.pdf](notes from Heike.pdf): scans of notes written by Heike for this project
- [README.md](README.md) and [README.Rmd](README.Rmd): this README and the R markdown document used to generate this README