### 2020-09-30 journals/00-objectives_and_ideas/00-objectives_and_ideas.Rmd
final clean up before making repo public

---
title: "Objectives and Ideas"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This journal explains the objectives and ideas for the research project on applying LIME to the Hamby bullet data that is detailed in this series of journals.

# Objectives

### Overall

- Understand how LIME works
- Consider ways to improvement LIME
- Apply LIME (or some improved version of LIME) to interpret the random forest model fit to the Hamby bullet data

### Ideas

**LIME diagnostics**

- add in feature selection methods to LIME input options
- think of a way to compute consistency across top two features
- Siggi suggests refitting the RF model to the perturbations and then continuing with LIME with the RF predictions from the new model - this may help to understand if the problems are due to the sampling procedure or LIME itself
- he also suggested looking into SMOTE for dealing the imbalance in the classes with sampling
- compare the simple models based on different number of bins using an F-test
- include a penalty for the number of parameters when choosing bins
- look at the AUC after binning
- compute a likelihood ratio prob TRUE / prob FALSE from the LIME ridge regression
- try visualizing the features from the test data using dimension reduction and
coloring them by variables suggested to be important by lime
- could try fitting a regression with interactions and see if LIME does a good job of explaining a model that is already interpretable
- come up with a test to compare between global and local explanations

**Explainer Models**

- read through papers giving an overview of the current explanation methods
- look into iml (https://www.youtube.com/watch?v=jP6Rg13PEkQ)
- look more into Molnar and his adviser Bernd
- look into [MAPLE](https://blog.ml.cmu.edu/2019/07/13/towards-interpretable-tree-ensembles/)

**Understanding LIME**

- Run a simulation to understand if LIME is working
  - could implement a couple of local linear dependencies
  - piece this together
  - could include interactions in the model
  - does lime find the local models?
- try fitting LASSO logistic model and leave one out approach (for multicollinarity)
- try reticulate R package to apply python version of lime
- look into literature on binning methods
- think about why R^2 would be better for some binning methods
- read new paper on Anchor

**Possible Improvements to LIME**

- determine the best number of bins to use for each variable
- try out subsampling idea

**Simulation with RF**

- think about how to implement a simulation with a random forest
- look back at 601 notes from Kaiser for model assessment
- look at Hadley's 'removing the blindfold' for ideas
- read book Dr. Dixon lent me on sensitivity analyses

**Random Forests Confidence Intervals**

- look into papers by Giles Hooker and Lucas Mentch
- Dr. Nettleton paper

**Random Forest Interpretations**

- read https://www.r-bloggers.com/explaining-predictions-random-forest-post-hoc-analysis-randomforestexplainer-package/

# Concerns

The following are some of the concerns that we have with the current state of the LIME algorithm.

- I'm nervous about the fact that the results can change due to the permutations. Is there a way to check for consistency? Does this only happen if you have correlated variables, or can it also happen with uncorrelated variables?
- When you have a large number of predictions to assess, would it be a good idea to focus in on the ones that have the best fitting linear model or produce the most consistent results?
- What can be done to improve the linear regression model fit? Maybe adjusting the number of bins or the kernel width would help with this.
- We think that the model explainer needs to be close enough to the complex model that it is trying to explain in order to do a good job of providing explanations. For example, the binned regression works okay with neural networks, but it does not work well with a random forest. Maybe a tree explainer or a logistic regression explainer would work better with a random forest.

# General Thoughts

- LIME is kind of like a jackknife technique